### 2020-06-16 TI_01_consistency.Rmd
consistency measures

---
title: "Triangulating iconicity: Coding, consistency & calibration"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='out/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

## Coding consistency
Code notebook for combining coding from two coders, working out consistency, and recording coding/recoding decisions.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","irr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```

## Data
Let's first load the original data and generate derivative XLS files to aid in the calibration.

```{r data}

d1 = read_excel("data\\ideophones_coded_MD_original.xlsx") %>% arrange(filename)
d2 = read_excel("data\\ideophones_coded_SP_original.xlsx") %>% arrange(filename)
d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge

d <- merge(d1,d2,by="filename")

# sanity check: are all ideophones in d1 and d2 also in d? (0 = yes)
d1$filename[d1$filename %notin% d$filename]
d2$filename[d2$filename %notin% d$filename]

d.F <- d %>%
  dplyr::select(c("filename","ideophone"),starts_with("F_")) %>%
  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")

d.M <- d %>%
  dplyr::select(c("filename","ideophone"),starts_with("M_")) %>%
  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")

# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
get_coding_categories <- function(data) {
  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
  gsub('_MD','',coder_1_names)
}

```

## Consistency
We compute coding consistency for the original coding to establish which categories of the coding scheme are usable as is and which need recalibration or reformulation. We use `irr` package to compute % agreement and Cohen's Kappa.

```{r irr}

# build functions to make it less repetitive
get_agree <- function(data,cat) {
  # takes a dataset and a coding category, looks up fields of that coding category, and computes agreement
  agree(dplyr::select(data,starts_with(cat)))$value
}
get_kappa <- function(data,cat) {
  # takes a dataset and a coding category, looks up fields of that coding category, and computes kappa
  kappa2(dplyr::select(data,starts_with(cat)))$value
}

# sanity check: does the function work? yes
get_agree(d.F,cat='F_redup')

# get list of coding categories and create dataframes for the consistency values
F_categories <- get_coding_categories(d.F)
F_consistency <- as.data.frame(F_categories) %>% dplyr::rename(category = F_categories)
M_categories <- get_coding_categories(d.M)
M_consistency <- as.data.frame(M_categories) %>% dplyr::rename(category = M_categories)

F_consistency$agree <- NA
F_consistency$kappa <- NA
for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
}
F_consistency
```

For Form, mean agreement in the original coding is **`r round(mean.na(F_consistency$agree))`%** and the lowest Kappa values are for F_intonation, F_voice, and F_partialredup, pointing to the need to resolve inconsistencies and sharpen those coding categories.

```{r irr_M}
M_consistency$agree <- NA
M_consistency$kappa <- NA
for (i in 1:(length(M_categories)-1)) {
  M_consistency$agree[i] <- (get_agree(data=d.M,cat=M_categories[i]))
  M_consistency$kappa[i] <- (get_kappa(data=d.M,cat=M_categories[i]))
}
M_consistency

```

For Meaning, mean agreement in the original coding is **`r round(mean.na(M_consistency$agree))`%** and the lowest Kappa values are for M-durative, M_weight, and M_irregular, pointing to the need to resolve inconsistencies and sharpen those coding categories.

## Outcomes
Using the consistency scores as a guide, we revisited the most problematic categories to (i) recalibrate coders' understanding of the categories, (ii) reformulate instructions and (iii) in some cases rename the coding category to reflect the revisions.

### Form
1.	*F_redupmod* was renamed (originally F_partialredup) and the coding instructions were sharpened to reflect the fact that we want to capture both partial and modified forms of reduplication.
2. *F_voice*: coding instructions were clarified to note that this question privileges consonants that crosslinguistically often enter into voicing oppositions (plosives and fricatives) as well as nasals, documented for their iconic uses across languages. This is close to Westermann’s use of this feature. We also considered a more continuous version of this feature that would weight or classify consonants based on their place on the sonority scale, but decided against it because the coding was already too fine-grained.
3. *F_finallength* was renamed (originally F_finalvowel) to capture word final lengthening of any final sound, whether final vowel or nasal, glide or trill. 
4.	*F_intonation*: coding instructions were clarified to include a decision procedure for forms with falling or rising intonation.

### Meaning
1. *M_irregularity*: coding instructions were modified to indicate that the primary semantic property here is one of irregularity in the sense of imbalance (a highly regular ridged pattern would not qualify)
2.	*M_weight*: coding instructions were sharpened to focus on binary oppositions of size, weight and brightness and to exclude Westermann’s hard/soft distinction, which we found orthogonal and conflicting in some cases. This also means that the middle value essentially means neutral or “not applicable”. It may be advisable to further break this up this feature, as there is no guarantee that size, weight and brightness pattern similarly across languages.
3. *M_durative*: coding instructions were sharpened to clarify that this targets events, not states and to clarify its relation to M_distribution.

The original coding sheet is available as X and the updated coding sheet as X_version_2.

### 2020-06-16 TI_02_congruence.Rmd
next step: congruence measures

--- 2020-06-16 TI_01_consistency.Rmd
@@
 ---
-title: "Triangulating iconicity: Coding, consistency & calibration"
+  title: "Triangulating iconicity: Using coding to predict experimental outcomes"
 author: "Mark Dingemanse & Stella Punselie"
@@
 
-## Coding consistency
-Code notebook for combining coding from two coders, working out consistency, and recording coding/recoding decisions.
+Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores.
 
@@
 # Packages
-list.of.packages <- c("tidyverse","readxl","writexl","irr")
+list.of.packages <- c("tidyverse","readxl","writexl")
 new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
@@
 }
-
-```
-
-## Consistency
-We compute coding consistency for the original coding to establish which categories of the coding scheme are usable as is and which need recalibration or reformulation.
-We use `irr` package to compute % agreement and Cohen's Kappa.
-
-```{r irr}
-
-# build functions to make it less repetitive
-get_agree <- function(data,cat) {
-  # takes a dataset and a coding category, looks up fields of that coding category, and computes agreement
-  agree(dplyr::select(data,starts_with(cat)))$value
-}
-get_kappa <- function(data,cat) {
-  # takes a dataset and a coding category, looks up fields of that coding category, and computes kappa
-  kappa2(dplyr::select(data,starts_with(cat)))$value
-}
-
-# sanity check: does the function work? yes
-get_agree(d.F,cat='F_redup')
-
-# get list of coding categories and create dataframes for the consistency values
-F_categories <- get_coding_categories(d.F)
-F_consistency <- as.data.frame(F_categories) %>% dplyr::rename(category = F_categories)
-M_categories <- get_coding_categories(d.M)
-M_consistency <- as.data.frame(M_categories) %>% dplyr::rename(category = M_categories)
-
-F_consistency$agree <- NA
-F_consistency$kappa <- NA
-for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
-  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
-  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
-}
-F_consistency
-```
-
-For Form, mean agreement in the original coding is **`r round(mean.na(F_consistency$agree))`%** and the lowest Kappa values are for F_intonation, F_voice, and F_partialredup, pointing to the need to resolve inconsistencies and sharpen those coding categories.
-
-```{r irr_M}
-M_consistency$agree <- NA
-M_consistency$kappa <- NA
-for (i in 1:(length(M_categories)-1)) {
-  M_consistency$agree[i] <- (get_agree(data=d.M,cat=M_categories[i]))
-  M_consistency$kappa[i] <- (get_kappa(data=d.M,cat=M_categories[i]))
-}
-M_consistency
-
-```
-
-For Meaning, mean agreement in the original coding is **`r round(mean.na(M_consistency$agree))`%** and the lowest Kappa values are for M-durative, M_weight, and M_irregular, pointing to the need to resolve inconsistencies and sharpen those coding categories.
-
-## Outcomes
-Using the consistency scores as a guide, we revisited the most problematic categories to (i) recalibrate coders' understanding of the categories, (ii) reformulate instructions and (iii) in some cases rename the coding category to reflect the revisions.
-
-### Form
-1.	*F_redupmod* was renamed (originally F_partialredup) and the coding instructions were sharpened to reflect the fact that we want to capture both partial and modified forms of reduplication.
-2.
-*F_voice*: coding instructions were clarified to note that this question privileges consonants that crosslinguistically often enter into voicing oppositions (plosives and fricatives) as well as nasals, documented for their iconic uses across languages.
-This is close to Westermann’s use of this feature.
-We also considered a more continuous version of this feature that would weight or classify consonants based on their place on the sonority scale, but decided against it because the coding was already too fine-grained.
-3. *F_finallength* was renamed (originally F_finalvowel) to capture word final lengthening of any final sound, whether final vowel or nasal, glide or trill. 
-4.	*F_intonation*: coding instructions were clarified to include a decision procedure for forms with falling or rising intonation.
-
-### Meaning
-1.
-*M_irregularity*: coding instructions were modified to indicate that the primary semantic property here is one of irregularity in the sense of imbalance (a highly regular ridged pattern would not qualify)
-2.	*M_weight*: coding instructions were sharpened to focus on binary oppositions of size, weight and brightness and to exclude Westermann’s hard/soft distinction, which we found orthogonal and conflicting in some cases.
-This also means that the middle value essentially means neutral or “not applicable”.
-It may be advisable to further break this up this feature, as there is no guarantee that size, weight and brightness pattern similarly across languages.
-3. *M_durative*: coding instructions were sharpened to clarify that this targets events, not states and to clarify its relation to M_distribution.
-
-The original coding sheet is available as X and the updated coding sheet as X_version_2.

### 2020-06-16 triangulating_iconicity.Rmd
next step: congruence measures

---
title: "Triangulating iconicity"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='out/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

## Predicting iconicity
Intro

## Setup

```{r preliminaries, results="hide"}
# Clear workspace
rm(list=ls())
# Packages
list.of.packages <- c("ggplot2","dplyr","tidyr","readxl","ggthemes","lme4","viridis","stringdist","VGAM")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# gg_jitterbox
source("gg_jitterbox.R")

```

## Data

```{r data}

d1 = read_excel("data\\ideophones_for_coding_new_MD.xlsx")
d2 = read_excel("data\\ideophones_for_coding_new_SP.xlsx")

# add logodds (it's safer to predict against logodds than raw proportion correct)
d <- d %>%
  mutate(cumulative_congruence = FM_congruent) %>%
  group_by(paper) %>%
  mutate(logodds = probit(score))

# convert congruency measures to factors
tofactors <- "language|category|paper|group|FM_congruent"
d[,grep(tofactors,names(d))] <- lapply(d[,grep(tofactors,names(d))], as.factor)

```

### 2020-06-16 TI_02_congruence.Rmd
congruency

@@
 # Packages
-list.of.packages <- c("tidyverse","readxl","writexl")
+list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis")
 new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
@@
 ## Data
-Let's first load the original data and generate derivative XLS files to aid in the calibration.
+Let's load the ground truth version of the coded data.
 
@@
 
-d1 = read_excel("data\\ideophones_coded_MD_original.xlsx") %>% arrange(filename)
-d2 = read_excel("data\\ideophones_coded_SP_original.xlsx") %>% arrange(filename)
-d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge
-
-d <- merge(d1,d2,by="filename")
-
-# sanity check: are all ideophones in d1 and d2 also in d? (0 = yes)
-d1$filename[d1$filename %notin% d$filename]
-d2$filename[d2$filename %notin% d$filename]
+d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)
 
 d.F <- d %>%
-  dplyr::select(c("filename","ideophone"),starts_with("F_")) %>%
+  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("F_")) %>%
   dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
-write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")
 
 d.M <- d %>%
-  dplyr::select(c("filename","ideophone"),starts_with("M_")) %>%
+  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("M_")) %>%
   dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
-write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")
 
-# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
-get_coding_categories <- function(data) {
-  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
-  gsub('_MD','',coder_1_names)
-}
+# list of the coding categories
+coding_categories <- d %>% dplyr::select(matches('_'),-matches('notes|meaning')) %>% names()
+
+```
+
+We add congruency measures that link F and M elements.
+
+```{r congruency_1}
+
+# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
+d <- d %>%
+  mutate(C_modality = ifelse(M_sound == 1,1,0),
+         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
+         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
+         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
+         C_durative = ifelse(F_finallength == 1 & M_durative == 1,1,0)) %>%
+  mutate(C_cumulative = C_modality + C_iterative + C_irregular + C_punctual)
+
+d %>%
+  group_by(C_cumulative) %>%
+    summarise(n=n(),mean.na(C_cumulative))
+
+```
+
+Let's add guessability from the Collabra and Language studies.
+
+```{r guessability}
+
+d.scores = read_excel("data\\ideophones_guessability.xlsx") 
+d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))
+
+```
+
+## Plots
+
+A quick first look. Needs fixing.
+
+```{r congruency_2}
+
+ggplot(data=d, aes(x=study,y=score)) +
+  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
+  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
+  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
+ggsave(filename = "figures//dotplot-all-blank.png",width=5,height = 7.5)
+
+# all data points by language
+ggplot(data=d, aes(x=study,y=score,fill=language,colour=language)) +
+  theme_tufte() + ylim(0,1) +
+  theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
+  geom_dotplot(colour="white",alpha=0.8,color=NA,stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
+ggsave(filename = "figures//dotplot-all-language.png",width=5,height = 7.5)
+
+# Westermann congruency: modality 
+ggplot(data=d, aes(x=study,y=score,fill=C_modality,colour=C_modality)) +
+  theme_tufte() + ylim(0,1) +
+  theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
+  scale_colour_manual(values=c("#C9C9C9","#ED7953")) +   scale_fill_manual(values=c("#C9C9C9","#ED7953")) +
+  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
+  geom_dotplot(colour="white",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
+ggsave(filename = "figures//dotplot-congruent_modality.png",width=5,height = 7.5)
+
+# Westermann congruency: cumulative
+ggplot(data=d, aes(x=study,y=score,colour=C_cumulative,fill=C_cumulative)) +
+  theme_tufte() + ylim(0,1) +
+  theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
+  geom_dotplot(colour="white",stackgroups = T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
+ggsave(filename = "out//dotplot-FM_congruent.png",width=5,height = 7.5)
+
+```

### 2020-06-16 TI_01_consistency.Rmd
calibration

@@
 The original coding sheet is available as X and the updated coding sheet as X_version_2.
+
+## Calibrated coding
+
+After reformulating and revising the coding scheme, we discussed all inconsistent cases to reach a consensus and recoded accordingly.
+Here we check whether we now indeed have a canonical, ground truth set of coded data.
+
+```{r consistency_2ndround}
+d1 = read_excel("data\\ideophones_coded_MD.xlsx") %>% arrange(filename)
+d2 = read_excel("data\\ideophones_coded_SP.xlsx") %>% arrange(filename)
+d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge
+
+d <- merge(d1,d2,by="filename")
+
+d.F <- d %>%
+  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("F_")) %>%
+  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
+
+d.M <- d %>%
+  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("M_")) %>%
+  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
+
+# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
+get_coding_categories <- function(data) {
+  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
+  gsub('_MD','',coder_1_names)
+}
+
+# get list of coding categories and create dataframes for the consistency values
+F_categories <- get_coding_categories(d.F)
+F_consistency <- as.data.frame(F_categories) %>% dplyr::rename(category = F_categories)
+M_categories <- get_coding_categories(d.M)
+M_consistency <- as.data.frame(M_categories) %>% dplyr::rename(category = M_categories)
+
+F_consistency$agree <- NA
+F_consistency$kappa <- NA
+for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
+  F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
+  F_consistency$kappa[i] <- (get_kappa(data=d.F,cat=F_categories[i]))
+}
+F_consistency
+
+```
+
+Hmm, a consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength.
+We can export only those columns to have a quick check.
+
+```{r inconsistency_output}
+
+inconsistent <- c('F_redup','F_monosyllabic','F_vowelquality','F_finallength')
+d.F.inconsistent <- d.F %>%
+  dplyr::select(filename,ideophone,meaning,meaning_NL,matches(paste(inconsistent,collapse="|"))) %>%
+  mutate(F_redup_C = ifelse(F_redup_MD == F_redup_SP,1,0),
+         F_monosyllabic_C = ifelse(F_monosyllabic_MD == F_monosyllabic_SP,1,0),
+         F_vowelquality_C = ifelse(F_vowelquality_MD == F_vowelquality_SP,1,0),
+         F_finallength_C = ifelse(F_finallength_MD == F_finallength_SP,1,0)) %>%
+  mutate(check = F_redup_C + F_monosyllabic_C + F_vowelquality_C + F_finallength_C) %>%
+  filter(check < 4)
+d.F.inconsistent <- d.F.inconsistent[, order(names(d.F.inconsistent))]
+
+write_xlsx(d.F.inconsistent,path="data\\ideophones_F_consistency_2ndround.xlsx")
+
+```
+
+Ultimately we want to output a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses.
+
+```{r write_final}
+
+# for now I'm taking just one coder
+d.coded <- read_excel("data\\ideophones_coded_SP.xlsx") %>% 
+  arrange(filename) %>%
+  dplyr::select(-matches('notes'))
+names(d.coded) <- gsub('_SP','',names(d.coded))
+
+notes <- d %>% dplyr::select(filename,matches('notes'))
+
+d.coded <- left_join(d.coded,notes)
+
+write_xlsx(d.coded,path="data\\ideophones_coded.xlsx")
+
+```

### 2020-06-17 TI_01_consistency.Rmd
more work on congruency

@@
 ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='out/',
+knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                       echo=TRUE, warning=FALSE, message=FALSE)
@@
 
-# sanity check: are all ideophones in d1 and d2 also in d? (0 = yes)
-d1$filename[d1$filename %notin% d$filename]
-d2$filename[d2$filename %notin% d$filename]
-
 d.F <- d %>%
@@
   dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
-write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")
+#write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")
 
@@
   dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
-write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")
+#write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")
 
@@
   coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
-  gsub('_MD','',coder_1_names)
+  gsub('MD','',coder_1_names)
 }
@@
 # sanity check: does the function work? yes
-get_agree(d.F,cat='F_redup')
+get_agree(d.F,cat='F_redup_')
 
@@
 
-# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
-get_coding_categories <- function(data) {
-  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
-  gsub('_MD','',coder_1_names)
-}
-
 # get list of coding categories and create dataframes for the consistency values
@@
 
-Ultimately we want to output a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses.
+* Ultimately we want to output a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses.
 

### 2020-06-17 TI_02_congruence.Rmd
update

--- 2020-06-16 TI_02_congruence.Rmd
@@
 ---
-  title: "Triangulating iconicity: Using coding to predict experimental outcomes"
+title: "Triangulating iconicity: From structure mapping to guessability"
 author: "Mark Dingemanse & Stella Punselie"
@@
   ```{r global_options, include=FALSE}
-knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='out/',
+knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                       echo=TRUE, warning=FALSE, message=FALSE)
@@
 # Packages
-list.of.packages <- c("tidyverse","readxl","writexl")
+list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","lme4","VGAM")
 new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
@@
 
+# custom functions
+source("TI_functions.R")
 ```
@@
 ## Data
-Let's first load the original data and generate derivative XLS files to aid in the calibration.
+Here we load the ground truth version of the coded data and add the independently collected guessability scores from the 2016 Collabra and Language papers.
 
@@
 
-d1 = read_excel("data\\ideophones_coded_MD_original.xlsx") %>% arrange(filename)
-d2 = read_excel("data\\ideophones_coded_SP_original.xlsx") %>% arrange(filename)
-d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge
+# get coding data
+d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)
 
-d <- merge(d1,d2,by="filename")
+coding_categories <- d %>% dplyr::select(matches('_'),-matches('notes|meaning')) %>% names()
 
-# sanity check: are all ideophones in d1 and d2 also in d? (0 = yes)
-d1$filename[d1$filename %notin% d$filename]
-d2$filename[d2$filename %notin% d$filename]
+# add guessability scores from the Collabra and Language studies.
 
-d.F <- d %>%
-  dplyr::select(c("filename","ideophone"),starts_with("F_")) %>%
-  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
-write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")
+d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
+  dplyr::select(-category)
+d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))
 
-d.M <- d %>%
-  dplyr::select(c("filename","ideophone"),starts_with("M_")) %>%
-  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
-write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")
+# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
+d <- d %>%
+  group_by(study) %>%
+  mutate(logodds = probitlink(score))
 
-# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
-get_coding_categories <- function(data) {
-  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
-  gsub('_MD','',coder_1_names)
-}
+```
+
+We add congruency measures that link F and M elements.
+
+* TO DISCUSS: We need to think about the evidential value of each of the congruency measures.
+Most apply to many words, which is good.
+If some of them only capture a few (as `C_durative` and `C_weight_tone` seem to do), they are not so interesting because they don't really capture recurring iconic mappings in the data.
+Or perhaps this may indicate a need to recode — e.g.
+perhaps `F_finallength` should not just map onto aspectual semantics (like `M_durative` but also to shape, sth like `M_long`: a very concrete kind of mapping.)
+
+```{r congruency_1}
+
+# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
+d <- d %>%
+  mutate(C_modality = ifelse(M_sound == 1,1,0),
+         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
+         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
+         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
+         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
+         C_durative = ifelse(F_finallength == 1 & M_durative == 1,1,0),
+         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight) == 2,1,0),
+         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
+         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
+         ) %>%
+  mutate(C_cumulative = C_modality + C_iterative + C_irregular + C_closure + C_punctual + C_durative + C_weight_voice + C_weight_vowel + C_weight_tone) %>%
+  mutate(C_any = ifelse(C_cumulative == 0,0,1),
+         C_atleast2 = ifelse(C_cumulative > 1,1,0),
+         C_ternary = ifelse(C_cumulative == 0,0,
+                            ifelse(C_cumulative == 1,1,2)))
+
+d %>%
+  group_by(C_cumulative) %>%
+    summarise(n=n())
+
+```
+
+## Plots
+
+A quick first look at some congruency measures. Uses a custom plotting function `icoplot()`, which is a wrapper for a ggplot object constructed with `geom_dotplot`.
+
+```{r icoplots}
+
+# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
+tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
+dp <- d
+dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
+
+# all data points by language
+icoplot()
+
+icoplot("C_modality")
+icoplot("C_iterative")
+icoplot("C_irregular")
+icoplot("C_punctual")
+icoplot("C_durative")
+icoplot("C_weight_voice")
+icoplot("C_weight_vowel")
+icoplot("C_weight_tone")
+
+icoplot("C_cumulative")
+
+icoplot("C_any")
+icoplot("C_atleast2")
+icoplot("C_ternary")
+
+# blank plot can be useful as background for slides
+ggplot(data=dp, aes(x=study,y=score)) +
+  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
+  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
+  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
+
+```
+
+## Exploring congruency blind spots
+What are we not capturing?
+Our coding and congruency measures are designed to be reproducible and human-codable.
+They are not exhaustive, and there may well be aspects that we are missing.
+Here we explore the congruency data to find highly guessable ideophones that our coding did not (yet) identify as having iconic mappings.
+
+```{r congruency_blindspots}
+
+d.blindspots <- d %>%
+  filter(C_cumulative == 0, score >0.5) %>%
+  dplyr::select(filename,ideophone,meaning,meaning_NL,language,score) %>%
+  arrange(-score)
+View(d.blindspots)
+
+```
+
+## Stats
+
+```{r congruency_stats_1}
+
+# simplest thing we can do is a correlation
+
+cor.test(d$logodds,d$C_cumulative)
+cor.test(d$score,d$C_cumulative)
+
+```

### 2020-06-17 TI_functions.R

@@ -1,0 +1,44 @@
+
+
+# icoplot() to make creating congruency plots easier
+# to do: build in variables for data source, score vs log odds, faceting by study, language, or semantic domain, etc.
+
+icoplot <- function(var=NULL,saveplot=FALSE) {
+  
+  if(is.null(var)) {
+    var <- "language"
+    print("No variable specified")
+  } 
+  print(paste("Plotting",var))
+  
+  # collect data
+  
+  # build plot
+  p <- ggplot(data=dp, aes_string(x="study",y="score",fill=substitute(var),colour=substitute(var))) +
+    theme_tufte() + ylim(0,1) + 
+    theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0))
+  # set colour scheme based on nature of variable
+  if (var %in% c("C_cumulative","C_ternary","language")) {
+    p <- p +
+      scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+      scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9)
+  } else {
+    p <- p +
+      scale_colour_manual(values=c("#B6308B","#FCCE25")) +   scale_fill_manual(values=c("#B6308B","#FCCE25"))
+  }
+  p <- p +
+    stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
+    geom_dotplot(colour="white",method="dotdensity",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center",na.rm=T)
+  
+  
+  # save plot
+  if (saveplot) {
+    filename <- paste0("dotplot-",var,".png")
+    print(paste("Saving as",filename))
+    ggsave(file=paste0("figures//",filename,width=5,height=7.5))
+  }
+  
+  # plot plot
+  suppressWarnings(print(p))
+  
+}

### 2020-06-17 TI_03_ratings.Rmd
next steps

--- 2020-06-16 TI_02_congruence.Rmd
@@
 ---
-  title: "Triangulating iconicity: Using coding to predict experimental outcomes"
+  title: "Triangulating iconicity: Iconicity ratings"
 author: "Mark Dingemanse & Stella Punselie"
@@
 ```
-
-## Data
-Let's first load the original data and generate derivative XLS files to aid in the calibration.
-
-```{r data}
-
-d1 = read_excel("data\\ideophones_coded_MD_original.xlsx") %>% arrange(filename)
-d2 = read_excel("data\\ideophones_coded_SP_original.xlsx") %>% arrange(filename)
-d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge
-
-d <- merge(d1,d2,by="filename")
-
-# sanity check: are all ideophones in d1 and d2 also in d? (0 = yes)
-d1$filename[d1$filename %notin% d$filename]
-d2$filename[d2$filename %notin% d$filename]
-
-d.F <- d %>%
-  dplyr::select(c("filename","ideophone"),starts_with("F_")) %>%
-  dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
-write_xlsx(d.F,path="data\\ideophones_F_consistency.xlsx")
-
-d.M <- d %>%
-  dplyr::select(c("filename","ideophone"),starts_with("M_")) %>%
-  dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
-write_xlsx(d.M,path="data\\ideophones_M_consistency.xlsx")
-
-# function to get the names of coding categories (note: this presupposes there is a coder _MD, change as needed)
-get_coding_categories <- function(data) {
-  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
-  gsub('_MD','',coder_1_names)
-}

### 2020-06-17 TI_04_triangulation.Rmd
next steps

--- 2020-06-17 TI_03_ratings.Rmd
@@
 ---
-  title: "Triangulating iconicity: Iconicity ratings"
+  title: "Triangulating iconicity: combining coding, guessing and rating measures"
 author: "Mark Dingemanse & Stella Punselie"

### 2020-06-26 TI_01_consistency.Rmd
coding scheme update

@@
 It may be advisable to further break this up this feature, as there is no guarantee that size, weight and brightness pattern similarly across languages.
-3. *M_durative*: coding instructions were sharpened to clarify that this targets events, not states and to clarify its relation to M_distribution.
+3.
+*M_long* was renamed (originally M_durative): coding instructions were broadened to capture percepts that are long in time and/or space (in line with M_distribution), and sharpened to clarify that for the temporal dimension it targets only events.
 

### 2020-06-26 TI_02_congruence.Rmd
tinkering with visualizations and gghalves

@@
 # Packages
-list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","lme4","VGAM")
+list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","gghalves","ggbeeswarm","viridis","lme4","VGAM")
 new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
@@
 
+# add Z score to account for differences in base success rate of Collabra and Language Studies
+d <- d %>%
+  group_by(study) %>%
+  mutate(score_z = scale(score,center=T,scale=T))
+
 ```
@@
 If some of them only capture a few (as `C_durative` and `C_weight_tone` seem to do), they are not so interesting because they don't really capture recurring iconic mappings in the data.
-Or perhaps this may indicate a need to recode — e.g.
-perhaps `F_finallength` should not just map onto aspectual semantics (like `M_durative` but also to shape, sth like `M_long`: a very concrete kind of mapping.)
+* Discussed: This made us realize a need to name `M_durative` to `M_long`, covering extent not just in time but also in space (a very concrete kind of mapping for `F_finallength`.
 
@@
          C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
-         C_durative = ifelse(F_finallength == 1 & M_durative == 1,1,0),
+         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
          C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight) == 2,1,0),
@@
          C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
-         ) %>%
-  mutate(C_cumulative = C_modality + C_iterative + C_irregular + C_closure + C_punctual + C_durative + C_weight_voice + C_weight_vowel + C_weight_tone) %>%
-  mutate(C_any = ifelse(C_cumulative == 0,0,1),
+         )
+
+# are some mappings mutually exclusive? then it might be slightly more elegant to combine them
+d %>%
+  mutate(C_event = C_iterative + C_punctual) %>%
+  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_event) %>%
+  filter(C_event > 1)
+
+d %>%
+  mutate(C_exclusive = C_long + C_closure) %>%
+  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_exclusive) %>%
+  filter(C_exclusive > 1)
+
+# add some cumulative iconicity scores
+d <- d %>%
+  mutate(C_cumulative = C_modality + C_iterative + C_irregular + C_closure + C_punctual + C_long + C_weight_voice + C_weight_vowel + C_weight_tone) %>%
+  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
          C_atleast2 = ifelse(C_cumulative > 1,1,0),
          C_ternary = ifelse(C_cumulative == 0,0,
-                            ifelse(C_cumulative == 1,1,2)))
+                            ifelse(C_cumulative == 1,1,
+                                   ifelse(C_cumulative == 2,2,3))))
+
+# cumulative scores can be further refined by considering only features that work for this sample — e.g., since C_irregular only adds noise, it may be excluded 
+d <- d %>%
+  mutate(C_simple = C_modality + C_iterative + C_closure + C_punctual + C_weight_tone + C_weight_vowel)
+
+# write this dataset for use in the next step
+write.csv(d,file="data/ideophones_coded_guessed.csv")
 
@@
   group_by(C_cumulative) %>%
-    summarise(n=n())
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_ternary) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_simple) %>%
+  summarise(n=n(),score=mean.na(score))
 
@@
 # all data points by language
-icoplot()
+icoplot() # icoplot() is loaded from TI_functions.R
 
@@
 icoplot("C_punctual")
-icoplot("C_durative")
+icoplot("C_long")
 icoplot("C_weight_voice")
@@
 
-icoplot("C_any")
+icoplot("C_any1")
 icoplot("C_atleast2")
@@
 
+### Tinkering with plots
+Just trying different ways of visualizing the data to find the most helpful and transparent one.
+
+```{r visuals}
+
+icoplot("C_simple")
+
+ggplot(data=dp, aes(x="all",y=score_z)) +
+  theme_tufte() +
+  geom_half_boxplot() +
+  geom_half_point(side="l") +
+  geom_half_violin(side="r") +
+  NULL
+
+ggplot(data=dp, aes(x=C_simple,y=logodds)) +
+  theme_tufte() +
+  geom_half_boxplot() +
+  geom_half_point() +
+  NULL
+ggplot(data=dp, aes(x=C_simple,y=score_z)) +
+  theme_tufte() +
+  geom_half_boxplot() +
+  geom_half_point() +
+  NULL
+
+ggplot(data=dp, aes(x=C_simple,y=logodds,color=C_simple)) +
+  theme_tufte() +
+  geom_half_boxplot() +
+  geom_half_dotplot() +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  NULL
+
+ggplot(data=dp, aes(x=C_ternary,y=logodds,color=C_ternary,fill=C_ternary)) +
+  theme_tufte() +
+  geom_half_violin() +
+  geom_half_point_panel() +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  facet_grid(~ study)
+
+ggplot(data=dp, aes(x=C_ternary,y=logodds,color=C_ternary,fill=C_ternary)) +
+  theme_tufte() +
+  geom_half_violin() +
+  geom_half_dotplot(dotsize=0.8) +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  facet_grid(~ study)
+
+ggplot(data=dp, aes(x=C_atleast2,y=logodds,color=C_atleast2,fill=C_atleast2)) +
+  theme_tufte() +
+  geom_half_violin() +
+  geom_half_dotplot(dotsize=0.4,stackratio=0.8,method="histodot") +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  facet_grid(~ study)
+
+ggplot(data=dp, aes(x=category,y=logodds,color=category)) +
+  theme_tufte() +
+  geom_half_violin() +
+  geom_half_point_panel() +
+  NULL
+
+ggplot(data=dp, aes(x=language,y=logodds,color=language)) +
+  theme_tufte() +
+  geom_half_violin() +
+  geom_half_point_panel() +
+  NULL
+
+ggplot(data=d, aes(x=language,y=C_ternary,color=language)) +
+  theme_tufte() +
+  geom_half_violin() +
+  geom_half_point_panel() +
+  NULL
+
+```
+
 ## Exploring congruency blind spots
@@
 They are not exhaustive, and there may well be aspects that we are missing.
-Here we explore the congruency data to find highly guessable ideophones that our coding did not (yet) identify as having iconic mappings.
+Some ideophones may also be guessed relatively high for reasons other than form-meaning analogies (e.g., they had a very unattractive foil or presented unforeseen confounds).
+Here we explore the congruency data to find highly guessed ideophones that our coding did not (yet) identify as having iconic mappings.
+We also look at the other end: ideophones coded as having an iconic mapping but not doing well in the binary forced choice task.
+
+If we sort by score most of the blindspots are from the Collabra study, which had a less forbidding design and therefore a higher baseline score.
 
@@
 
-d.blindspots <- d %>%
-  filter(C_cumulative == 0, score >0.5) %>%
-  dplyr::select(filename,ideophone,meaning,meaning_NL,language,score) %>%
-  arrange(-score)
-View(d.blindspots)
+d %>%
+  filter(C_cumulative == 0, logodds >0) %>%
+  dplyr::select(filename,ideophone,meaning,meaning_NL,language,logodds,score) %>%
+  arrange(-score) %>% ungroup %>%
+  slice(1:10)
+```
+
+```{r congruency_underperformers}
+# the reverse is also interesting: which ideophones were coded as having iconic mappings yet were not guessed greatly?
+
+d %>%
+  filter(C_cumulative > 0, logodds < -0.25) %>%
+  dplyr::select(filename,ideophone,meaning,meaning_NL,C_cumulative,language,logodds,score) %>%
+  arrange(score) %>% ungroup() %>%
+  slice(1:10)
 
@@
 ## Stats
+As the visualizations already make clear there is a non-trivial positive relation between the presence of iconic mappings (in a cumulative sense) and the guessability score determined in experimental work.
 
@@
 cor.test(d$logodds,d$C_cumulative)
-cor.test(d$score,d$C_cumulative)
+cor.test(d$score_z,d$C_cumulative)
 

### 2020-06-26 TI_03_ratings.Rmd
tinkering with visualizations and gghalves

@@
 
-Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores.
+Code notebook for a study of the relation between iconicity ratings and experimentally collected guessability scores.
 
@@
 # Packages
-list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis")
+list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","lme4","VGAM")
 new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
@@
 ```
+
+## Data
+Here we load the combined coding and guessability data and then add the rating data.
+
+```{r data}
+# get coding data
+d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)
+
+coding_categories <- d %>% dplyr::select(matches('_'),-matches('notes|meaning')) %>% names()
+
+# add guessability scores from the Collabra and Language studies.
+
+d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
+  dplyr::select(-category)
+d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))
+
+# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
+d <- d %>%
+  group_by(study) %>%
+  mutate(logodds = probitlink(score))
+
+# add mean ratings
+d.ratings = read_excel("data\\ideophones_rated_means.xlsx") %>%
+  dplyr::select(-c(item,category,list))
+d <- left_join(d,d.ratings,by=c("filename","language","study"))
+
+# convert ratings to z-scores for comparison with guessability (logodd) scores
+d$rating_z <- scale(d$rating,center=T,scale=T)
+
+```
+
+## Statistics
+```{r stats}
+# before calculating the correlation coefficient, we visualise the correlation between the guessing and rating data
+ggplot(d,aes(x=rating_z,y=logodds)) + geom_point(colour="dark grey",position="jitter") + geom_smooth(method=lm,colour="black")
+ggsave("figures\\TI_03_corr.png",height=5,width=7.5)
+
+ggplot(d,aes(x=rating_z,y=logodds)) + geom_point(aes(color=category,shape=language),position="jitter")
+ggsave("figures\\TI_03_corr_bylanguageandcategory.png",height=5,width=7.5)
+
+# Pearson correlation
+cor.test(d$logodds,d$rating_z,method="pearson")
+
+```
+
+Pearson's correlation coefficient indicates that there is a strong positive correlation between the guessing and rating scores, r(237) = .57, p < .001.

### 2020-06-26 TI_functions.R

@@ -18,7 +18,7 @@ icoplot <- function(var=NULL,saveplot=FALSE) {
     theme_tufte() + ylim(0,1) + 
     theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0))
   # set colour scheme based on nature of variable
-  if (var %in% c("C_cumulative","C_ternary","language")) {
+  if (var %in% c("C_cumulative","C_ternary","language","C_simple")) {
     p <- p +
       scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
       scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9)

### 2021-06-09 TI_03_ratings.Rmd
cleaning up and refactoring for release

--- 2020-06-17 TI_03_ratings.Rmd
@@
 # Packages
-list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis")
+list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","gghalves","plotly","lme4","VGAM","plyr","emmeans","car")
 new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
@@
 ```
+
+## Rating: descriptives
+
+```{r}
+# load dataframe with all ratings
+read.csv("data\\ideophones_rated_uncorr.csv",header=TRUE,na.strings=c("", "NA"),sep=";",check.names=FALSE) -> d.uncorr
+
+# to long format
+d.uncorr <- gather(d.uncorr, item, rating, 6:245)
+d.uncorr <- d.uncorr[,c(1,6,7)]
+colnames(d.uncorr) <- c("rater","item","rating")
+#remove empty cells
+d.uncorr <- subset(d.uncorr, !is.na(d.uncorr[,"rating"]))
+#change ratings to numeric values with decimals
+as.numeric(sub(",", ".", d.uncorr$rating, fixed = TRUE)) -> d.uncorr$rating
+
+# check for inconsistent raters with person-total correlation based on Motamedi et al. (2019)
+ptc <- d.uncorr[,c("item","rater","rating")]
+personTotalCorrelationCorrected <- function(ptc)
+  {
+  require(tidyverse)
+  ptc %>%
+    dplyr::rename(raterFocal = rater,
+           ratingFocal = rating) %>%
+    full_join(ptc, by='item') %>%
+    filter(rater!=raterFocal) %>%
+    group_by(item, raterFocal, ratingFocal) %>%
+    summarise(ratingsOthers = mean(rating, na.rm=TRUE)) %>%
+    base::split(.$raterFocal) %>%
+    map(~cor(.$ratingFocal, .$ratingsOthers, 
+             use="pairwise.complete.obs")) %>%
+    as.data.frame %>%
+    gather %>%
+    dplyr::rename(rater = key, perTotCor = value)
+}
+View(personTotalCorrelationCorrected(ptc))
+
+#plot person-total correlations
+ptc %>%
+select(item, rater, rating) %>%
+personTotalCorrelationCorrected %>%
+ggplot(aes(x=perTotCor)) + 
+  geom_density() + 
+  geom_histogram(stat="bin", 
+                 alpha=0.4,
+                 aes(y=..density..),
+                 binwidth=0.1) +
+  labs(x="Person-total correlation")
+ggsave("figures\\rating_PerTotCorr.png",height=5,width=7.5)
+```
+
+One participant (N045) showed a negative person-total correlation, which according to Curran (2016) indicates a careless responder rendering invalid data.
+Inspection of this participant’s data showed they had indeed given the same answer to every question.
+I therefore decided to remove this participant from the dataset.
+
+Two other participatns (N022 and N034) also had to be removed because of their knowledge of Japanese and Korean, resulting in a total of 75 participants to be included in the analysis.
+
+```{r}
+# load corrected data
+d <- read_xlsx("data\\ideophones_rated.xlsx")
+d.means <- read_xlsx("data\\ideophones_rated_means.xlsx")
+
+# get means and standard deviations across all ideophones and for each language and category
+mean(d$rating)
+  # 2.948339
+sd(d$rating)
+  # 1.297106
+sumcat <- ddply(d,~category,summarise,mean=mean(rating),sd=sd(rating))
+sumlang <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))
+
+```
+
+Because the category "Other" is quite a vague residual category with only 17 items (as opposed to 30-40 items for the other categories), I decided to remove this category for now.
+I will keep a copy that includes the ratings for "Other".
+
+```{r}
+#create copy of data including the category "Other"
+d.other <- read_xlsx("data\\ideophones_rated.xlsx")
+d.m.other <- read_xlsx("data\\ideophones_rated_means.xlsx")
+
+#remove category "Other" from d and d.means
+d <- d[!grepl("Other",d$category),]
+d.means <- d.means[!grepl("Other",d.means$category),]
+
+# make new summary for language (without the Japanese "other" ideophones)
+sumlang2 <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))
+```
+
+Here we plot the distribution of the rating data.
+
+```{r}
+#convert some columns to factors
+cols <- c("category","language","study")
+d[cols] <- lapply(d[cols], factor)
+d.means[cols] <- lapply(d.means[cols], factor)
+
+#combined boxplot and scatterplot by category
+ggplot(d.means, aes(x=reorder(category,-rating),y=rating,color=category)) +
+  theme_tufte(base_size=16) +
+  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
+  geom_half_point(show.legend=F) +
+  scale_colour_viridis_d(option="D",alpha=0.8) +
+  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
+  "Texture","Colour/Visual"))
+ggsave("figures\\rating_category.png",height=5,width=7.5)
+
+#same, but by language
+ggplot(d.means, aes(x=reorder(language,-rating),y=rating,color=language)) +
+  theme_tufte(base_size=16) +
+  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
+  geom_half_point(show.legend=F) +
+  scale_colour_viridis_d(option="D",alpha=0.8) +
+  xlab("category")
+ggsave("figures\\rating_language.png",height=5,width=7.5)
+
+```
+
+## Statistics
+
+Now we want to see which categories and languages differ significantly from the others in their iconicity ratings.
+
+First, we check the assumptions.
+
+```{r}
+# check normality by visual inspection
+ggqqplot(d, "rating", facet.by = "category")
+  # all points fall approximately along the reference line, so normality can be assumed
+
+# check homogeneity of variance with Levene's test
+leveneTest(rating~category,data=d)
+leveneTest(rating~language,data=d)
+  # both are not significant (> .05), so homogeneity of variance can be assumed
+```
+
+Now we construct a linear model with category as a predictor of the iconicity ratings.
+
+```{r}
+summary(lm.cat <- lm(rating ~ category, d))
+
+# pairwise comparisons with Tukey adjustments in the emmeans package
+emm.cat <- emmeans(lm.cat, "category")
+pairs(emm.cat)
+pwpm(emm.cat)
+```
+
+And here we do the same with language as predictor.
+
+```{r}
+summary(lm.lang <- lm(rating ~ language, d))
+
+# pairwise comparisons in emmeans package
+emm.lang <- emmeans(lm.lang, "language")
+pairs(emm.lang)
+pwpm(emm.lang)
+```
+
+## Combined data
+Here we load the combined coding and guessability data and then add the rating data.
+
+```{r}
+# get coded and guessed data from previous step (TI_02_congruence)
+d = read.csv("data\\ideophones_coded_guessed.csv")
+
+coding_categories <- d %>% dplyr::select(matches('F_|M_'),-matches('notes|meaning')) %>% names()
+
+# add mean ratings
+d.ratings = read_excel("data\\ideophones_rated_means.xlsx") %>%
+  dplyr::select(-c(item,category,list))
+d <- left_join(d,d.ratings,by=c("filename","language","study"))
+
+# convert ratings to z-scores for comparison with guessability (logodd) scores
+d$rating_z <- scale(d$rating,center=T,scale=T)
+
+#remove category "Other"
+d <- d[!grepl("Other",d$category),]
+
+```
+
+## Visualisations
+Let's first visualise the relation between the rating and guessing data.
+
+```{r}
+ggplot(d,aes(x=rating_z,y=score_z)) +   
+  geom_point(aes(colour=study),position="jitter") +
+  geom_smooth(method=lm,colour="black") +
+  theme_tufte(base_size = 16) +
+  xlab("rating") + ylab("guessability")
+ggsave("figures\\rating_guessing_corr.png",height=5,width=7.5)
+
+ggplot(d,aes(x=rating_z,y=score_z)) +   
+  geom_point(aes(colour=category,shape=study),position="jitter") +
+  geom_smooth(method=loess,colour="black") +
+  theme_tufte(base_size = 16) +
+  xlab("rating") + ylab("guessability")
+```
+
+There appears to be a positive correlation between the rating and guessing data.
+
+Now let's visualise the relations between the coding, guessing and rating data.
+
+```{r}
+
+ggplot(d,aes(x=rating_z,y=logodds, color=C_cumulative)) +
+  geom_point() +
+  geom_smooth(method=loess) +
+  scale_fill_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
+  NULL
+
+plot_ly(data=d,x=rating_z,y=score_z,color=C_simple)
+
+```
+
+## Statistics
+```{r}
+# Pearson correlation
+cor.test(d$score_z,d$rating_z,method="pearson")
+cor.rank <- cor.test(d$score_z,d$C_cumulative,method="spearman")
+```
+
+Pearson's correlation coefficient indicates that there is indeed a strong positive correlation between the guessing and rating data, r(220) = .61, p < .001.
+
+```{r}
+#check where the guessing and rating data deviate
+d %>%
+  filter(rating_z-score_z > 2 | rating_z-score_z < -2)%>%
+  dplyr::select(filename,ideophone,rating,rating_z,language,logodds,score,score_z) %>%
+  arrange(score) %>% ungroup()
+```

### 2021-06-09 TI_03_ratings_MD.Rmd
cleaning up and refactoring for release

---
  title: "Triangulating iconicity: Iconicity ratings"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

  ```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

Code notebook for a study of the relation between iconicity ratings and experimentally collected guessability scores.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","gghalves","plotly","lme4","VGAM")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```

## Data
Here we load the combined coding and guessability data and then add the rating data.

```{r data}
# get coded and guessed data from previous step
d = read.csv("data\\ideophones_coded_guessed.csv")

coding_categories <- d %>% dplyr::select(matches('F_|M_'),-matches('notes|meaning')) %>% names()

# add mean ratings
d.ratings = read_excel("data\\ideophones_rated_means.xlsx") %>%
  dplyr::select(-c(item,category,list))
d <- left_join(d,d.ratings,by=c("filename","language","study"))

# convert ratings to z-scores for comparison with guessability (logodd) scores
d$rating_z <- scale(d$rating,center=T,scale=T)[,1]

# write file for use in next step
write.csv(d, file="data\\ideophones_coded_guessed_rated.csv")

```

## Statistics
```{r stats}
# before calculating the correlation coefficient, we visualise the correlation between the guessing and rating data
ggplot(d,aes(x=rating_z,y=logodds)) + 
  geom_point(colour="dark grey",position="jitter") + 
  geom_smooth(method=lm,colour="black")
ggsave("figures\\TI_03_corr.png",height=5,width=7.5)

ggplot(d,aes(x=rating_z,y=logodds)) + 
  geom_point(aes(color=category,shape=language),position="jitter")
ggsave("figures\\TI_03_corr_bylanguageandcategory.png",height=5,width=7.5)

# Pearson correlation
cor.test(d$logodds,d$rating_z,method="pearson")

```

Pearson's correlation coefficient indicates that there is a strong positive correlation between the guessing and rating scores, r(237) = .57, p < .001.

### 2021-06-09 paper_coding_1_consistency.Rmd
cleaning up and refactoring for release

--- 2020-06-26 TI_01_consistency.Rmd
@@
 F_consistency
+mean.na(F_consistency$kappa)
+
 ```
@@
 
-For Meaning, mean agreement in the original coding is **`r round(mean.na(M_consistency$agree))`%** and the lowest Kappa values are for M-durative, M_weight, and M_irregular, pointing to the need to resolve inconsistencies and sharpen those coding categories.
+For Meaning, mean agreement in the original coding is **`r round(mean.na(M_consistency$agree))`%** and the lowest Kappa values are for M_durative, M_weight, and M_irregular, pointing to the need to resolve inconsistencies and sharpen those coding categories.
 
@@
 
-Hmm, a consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength.
+A consistency check at this point shows we're nearly there but we're not 100% consistent on F_redup, F_monosyllabic and F_vowelquality and F_finallength.
 We can export only those columns to have a quick check.
+However, as we do not aim for 100% consistency, this is more for informative purposes.
 
@@
 
-* Ultimately we want to output a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses.
+### Consensus version
+The final result of the coding consistency is a consensus or ground truth file `ideophones_coded.xlsx`, to be used in the next step of the analyses.
+We do this this by taking the (calibrated and revised) data from one coder.
 
-```{r write_final}
+```{r create_consensus_version}
 
-# for now I'm taking just one coder
 d.coded <- read_excel("data\\ideophones_coded_SP.xlsx") %>% 
@@
 
-write_xlsx(d.coded,path="data\\ideophones_coded.xlsx")
+```
+
+## Adding congruency measures
+We add congruency measures by combining F and M features as described in the coding scheme.
+
+```{r congruency_1}
+
+# note that we overwrite the old d with the new consensus version here
+
+# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
+d <- d.coded %>%
+  mutate(C_modality = ifelse(M_sound == 1,1,0),
+         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
+         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
+         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
+         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
+         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
+         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight == 2),1,0),
+         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
+         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
+         )
+
+# some mappings are mutually exclusive, so we combine them
+
+#sanity check: are they indeed mutually exclusive? these calls should come up empty
+d %>% 
+  mutate(C_aspect = C_iterative + C_punctual) %>%
+  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_aspect) %>%
+  filter(C_aspect > 1)
+d %>%
+  mutate(C_length = C_long + C_closure) %>%
+  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) %>%
+  filter(C_length > 1)
+
+d <- d %>%
+  mutate(C_aspect = C_iterative + C_punctual,
+         C_length = C_long + C_closure)
+
+# C_weight_ subfeatures are linked to one and the same meaning feature M_weight so they are recombined into C_magnitude
+# (note: this may be subject to revision, as the subfeatures may not deserve equal weight)
+
+d <- d %>%
+  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
+         C_magnitude = ifelse(C_weight == 0,0,1)) # i.e. a magnitude value that is 1 if any weight feature was congruent
+
+# add cumulative iconicity scores (and some others)
+# any1 = 1 if an ideophone has at least 1 congruency feature of any type
+# atleast2 = if an ideophone has at least 2 congruency features of any type
+# ternary = simple way of conflating the (smaller) upper regions: 1=1, 2=2, >2 = 3
+d <- d %>%
+  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) %>%
+  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
+         C_atleast2 = ifelse(C_cumulative > 1,1,0),
+         C_ternary = ifelse(C_cumulative == 0,0,
+                            ifelse(C_cumulative == 1,1,
+                                   ifelse(C_cumulative == 2,2,3)))) # i.e. 1, 2 or 3
+
+# we can have a first look at the basic distribution of guessability by iconicity
+d %>%
+  group_by(C_cumulative) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_any1) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_atleast2) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_ternary) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_simple) %>%
+  summarise(n=n(),score=mean.na(score))
+
+# finally write this to XLS and CSV
+write_xlsx(d,path="data\\ideophones_coded.xlsx")
+write.csv(d,file="data\\ideophones_coded.csv")
 

### 2021-06-09 paper_coding_2_analysis.Rmd
cleaning up and refactoring for release

---
title: "Triangulating iconicity: Coding analysis"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

  ```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","gghalves","ggbeeswarm","viridis","lme4","VGAM","cowplot")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

# custom functions
source("TI_functions.R")
```

## Data
Here we load the ground truth version of the coded data and add the independently collected guessability scores from the 2016 Collabra and Language papers.

```{r data}

# get consensus coding data
d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)

# add guessability scores from the Collabra and Language studies.

d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
  dplyr::select(-category)
d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))

# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
d <- d %>%
  group_by(study) %>%
  mutate(logodds = probitlink(score))

# add Z score to make scores more comparable in plots across studies
d <- d %>%
  group_by(study) %>%
  mutate(score_z = scale(score,center=T,scale=T))

# get ratings data
d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
  dplyr::select(-category,-list,-item)
d <- left_join(d,d.ratings,by=c("filename","study","language"))

# add z score for ratings
d$rating_z <- scale(d$rating,center=T,scale=T)

# write this dataset as the fullest dataset
write.csv(d,file="data/ideophones_coded_guessed_rated.csv")
write_xlsx(d,path="data/ideophones_coded_guessed_rated.xlsx")

# also write a simplified copy for easy inspection in Excel (helpful when writing)
d %>%
  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,rating,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
  write_xlsx(path="data/ideophones_coded_guessed_rated_simplified.xlsx")

```

## Figures
Here we generate some descriptive stats and create the main figures for the paper.

```{r descriptive_stats}

d %>%
  group_by(study) %>%
  summarize(max=max(score),min=min(score),mean=mean.na(score))

```

### Figure: Overview panel
The overview panel is created separately but we use the data from the guessability studies as a canvas.

```{r figures_paper_guessability_canvas}

# blank plot can be useful as a canvas 
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="black",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

ggsave("figures\\paper-guessability-blank.pdf",height=5,width=7.5)

d %>% filter(filename=="Korean_23_Motion_org") %>% View()

```

### Figure: Guessability by study and iconicity rating

```{r figures_paper_guessability_rating}
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=study,colour=study)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
  NULL

pB <- ggplot(data=dp, aes(x=rating,y=score,fill=study,colour=study)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") +
  labs(title="Guessability (both studies) by iconicity rating",
       x="iconicity rating",
       y="") +
  geom_point(size=2.0,shape=21) +
  geom_smooth(method="lm") +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)

```

### Figure: Guessability and cumulative iconicity

```{r figures_paper_guessability_coding}

# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

# panel A: scores by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by study",
       x="study",
       y="guessability (% correct)") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
  NULL

pB <- ggplot(data=dp, aes(x=C_cumulative,y=score_z,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Guessability (both studies) by cumulative iconicity",
       x="cumulative iconicity",
       y=expression('guessability ('~italic(z)~')')) +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB.png",height=5,width=9)

```

### Figure: Ratings and cumulative iconicity

```{r ratings_and_C}

# panel A: ratings by study and cumulative iconicity
pA <- ggplot(data=dp, aes(x="",y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="all ideophones",
       y="mean iconicity rating") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  geom_beeswarm(cex=2.5) +
  NULL

pB <- ggplot(data=dp, aes(x=C_cumulative,y=rating,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
  labs(title="... by independently coded structural correspondences",
       x="cumulative iconicity",
       y="mean iconicity rating") +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
ggsave("figures\\paper_panelAB_ratings.png",height=5,width=9)

```

### Figure: Mappings and semantic domains

```{r figure_mappings_domains}

library(forcats)
dp.categories <- d %>%
  filter(category %notin% "Other") %>%
  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  ungroup() %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
t.test(group1,group2)
library(effsize)
cohen.d(group1,group2)

# also when we exclude C_modality? yes
dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)
dp.categories %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
t.test(group1,group2)
cohen.d(group1,group2)

ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Cumulative iconicity by semantic domain") +
  geom_half_violin() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Cumulative iconicity by semantic domain") +
  geom_hex() +
  NULL

```

### Raincloud plots
While the icoplots provide a useful first visualisation, for in a paper we would probably want to use something like raincloud plots that show key features of the distribution across groups.

```{r visuals}

# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

# panel A: scores by study and cumulative iconicity
ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank()) + 
  labs(title="Guessability by study and cumulative iconicity") +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

ggplot(data=dp, aes(x=C_cumulative,y=logodds)) +
  theme_tufte() +
  labs(title="Guessability (log odds) by cumulative iconicity",
       x="cumulative iconicity",
       y="guessability (log odds)") +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

ggplot(data=dp, aes(x=C_any1,y=logodds)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

ggplot(data=dp, aes(x=C_atleast2,y=logodds)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

ggplot(data=dp, aes(x=C_ternary,y=logodds)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

ggplot(data=dp, aes(x="all",y=score_z)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point(side="l") +
  geom_half_violin(side="r") +
  NULL

ggplot(data=dp, aes(x=C_cumulative,y=logodds,color=C_cumulative,fill=C_cumulative)) +
  theme_tufte() +
  geom_half_violin() +
  geom_half_point_panel() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9)

icoplot("C_cumulative")

ggplot(data=dp, aes(x=C_cumulative,y=logodds,color=C_cumulative,fill=C_cumulative)) +
  theme_tufte() +
  geom_half_violin() +
  geom_half_point_panel() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  facet_grid(~ study)

```

### Dotplots

A quick look at some congruency measures. Uses a custom plotting function `icoplot()`, which is a wrapper for a ggplot object constructed with `geom_dotplot`.

```{r icoplots}

# all data points by language
icoplot() # icoplot() is loaded from TI_functions.R

icoplot("C_modality")
icoplot("C_aspect")
icoplot("C_length")
icoplot("C_magnitude")
icoplot("C_irregular")
icoplot("C_cumulative")

# how do the constituents of C_magnitude fare?
icoplot("C_magnitude")
icoplot("C_weight_voice")
icoplot("C_weight_vowel")
icoplot("C_weight_tone")

icoplot("C_any1")
icoplot("C_atleast2")
icoplot("C_ternary")
icoplot("C_simple")

# blank plot can be useful as background for slides
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

# facet by language to discover crosslinguistic diversity
ggplot(data=dp, aes(x=study,y=score)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")

```

## Numbers

### Coding and guessability
As the visualizations already make clear there is a non-trivial positive relation between the presence of iconic mappings (in a cumulative sense) and the guessability score determined in experimental work.

The simplest thing to do is a simple correlation. Looks like there is an r=0.30 correlation between logodds and C_cumulative.

```{r congruency_stats_1}

# simplest thing we can do is a correlation.

cor.test(d$logodds,d$C_cumulative,method="pearson")

d.Collabra <- d %>% filter(study == "Collabra")
d.Language <- d %>% filter(study == "Language")
cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")
cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")

p0 <- cor.test(d$logodds,d$C_cumulative,method="pearson")$p.value
p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")$p.value
p2 <- cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")$p.value
p.adjust(c(p0,p1,p2),method="bonferroni")

```

But it's probably more useful to predict logodds score on the basis of what we know about paper (= a proxy for method) and the main measures of form-meaning congruence. For this we can use logistic regression.

A null model `m0` predices logodds by paper, controlling for item-level differences. A model `m1` with cumulative congruence as an additional fixed effect fares significantly better.

For the fixed effects, m1 lists a 0.182 improvement in logodds for C_cumulative. According to my interpretation this means that every added iconic feature leads to a ~7% boost in proportion correct.

Also note that, accounting for possible difference in coded cumulative iconicity across studies, the Language paper has a negative fixed effect of -0.524, which translates to a 19% guessability penalty for the relatively unforgiving method of the Language study (or alternatively, a similar-sized boost for the methods of the Collabra study).

```{r congruency_stats_2}

m0 <- lmer(logodds ~ study + (1|ideophone), data=d)
m1 <- lmer(logodds ~ study + C_cumulative + (1|ideophone), data=d)
anova(m0,m1)

fixef(m1)

probitlink(0.1823,inverse=T)-0.5
probitlink(-0.5242,inverse=T)-0.5

```

We can assess the relative weight of different features using model comparison. We build models by adding the congruency measures in rank order of attestation in the coded data. Doing this we find that C_aspect, C_modality and C_length are responsible for the bulk of the variance. Compared to the null model that only represents the baseline difference between studies, model `m8` fares significantly better (log likelihood difference = 21.3).

The model estimates given in the output of `fixef(m8)` represent the contribution of a given fixed effect while controlling for others, but they are NOT additive (i.e. can't be summed to arrive at the effect of `C_cumulative`).

* `C_aspect`: 0.181 logodds (+7% boost)
* `C_modality`: 0.177 logodds (+7% boost)
* `C_length`: 0.443 logodds (+17% boost)

```{r congruency_stats_3}

m3 <- lmer(logodds ~ study + C_aspect + (1|ideophone), data=d)
m4 <- lmer(logodds ~ study + C_aspect + C_magnitude + (1|ideophone), data=d)
m5 <- lmer(logodds ~ study + C_aspect + C_magnitude + C_modality + (1|ideophone), data=d)
m6 <- lmer(logodds ~ study + C_aspect + C_magnitude + C_modality + C_length + (1|ideophone), data=d)
m7 <- lmer(logodds ~ study + C_aspect + C_magnitude + C_modality + C_length + C_irregular + (1|ideophone), data=d)
m8 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)

anova(m0,m3) # aspect helps
anova(m3,m4) # magnitude doesn't additionally help
anova(m4,m5) # modality helps
anova(m3,m5) # modality helps also relative to m3
anova(m5,m6) # length helps
anova(m6,m7) # irregular doesn't additionally help
anova(m5,m8) # aspect + modality + length is where it's at
# so m8 is the best performing most complex model
anova(m0,m8)
summary(m8)
fixef(m8)

probitlink(fixef(m8)[1],inverse=T)-0.5
probitlink(fixef(m8)[2],inverse=T)-0.5
probitlink(fixef(m8)[3],inverse=T)-0.5
probitlink(fixef(m8)[4],inverse=T)-0.5
probitlink(fixef(m8)[5],inverse=T)-0.5

```

### Semantic domains
Our prediction is that Sound and Motion ideophones should show more structural correspondences than the other domains. This is borne out.

```{r coding_semanticdomains}

library(forcats)
dp.categories <- d %>%
  filter(category %notin% "Other") %>%
  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

dp.categories %>%
  ungroup() %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
t.test(group1,group2)
library(effsize)
cohen.d(group1,group2)

# also when we exclude C_modality? yes
dp.categories %>%
  group_by(category) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)
dp.categories %>%
  group_by(category_binary) %>%
  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
  arrange(-mean_C)

group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
t.test(group1,group2)
cohen.d(group1,group2)

```

We can also look in more detail at the relative differences by semantic domain by building a mixed model that predicts cumulative iconicity with domain as a predictor.

```{r coding_category_lme}

m0 <- lm(C_cumulative ~ category, data=d)
effsizes(m0)

```

Another prediction is that Sound and Motion ideophones should receive higher iconicity ratings than other domains. This prediction is also borne out.

```{r rating_domains}

group1 <- subset(dp.categories, category_binary == "SoundMotion")$rating
group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$rating
t.test(group1,group2)
cohen.d(group1,group2)

```

### Coding and ratings

As above, the simplest thing to do is run a simple correlation: there is an r=0.43 correlation between meaning iconicity ratings and cumulative iconicity.

```{r coding_ratings}

cor.test(d$rating_z,d$C_cumulative,method="pearson")

```

As above, we can go a bit deeper by using mixed effects modelling. We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away. However, the studies are also different in terms of

```{r coding_ratings_lme}

m0 <- lmer(rating ~ study + (1|ideophone), data=d)
m1 <- lmer(rating ~ study + C_cumulative + (1|ideophone), data=d)
anova(m0,m1)

fixef(m1)

```

We can assess the relative weight of different features using model comparison. We build models by adding the congruency measures in rank order of attestation in the coded data. Doing this we find that C_aspect, C_modality and C_length are responsible for the bulk of the variance. Compared to the null model that only represents the baseline difference between studies, model `m8` fares significantly better (log likelihood difference = 21.3).

The model estimates given in the output of `fixef(m8)` represent the contribution of a given fixed effect while controlling for others, but they are NOT additive (i.e. can't be summed to arrive at the effect of `C_cumulative`).

* `C_aspect`: 0.181 logodds (+7% boost)
* `C_modality`: 0.177 logodds (+7% boost)
* `C_length`: 0.443 logodds (+17% boost)

```{r congruency_stats_3}

m3 <- lmer(rating ~ study + C_aspect + (1|ideophone), data=d)
m4 <- lmer(rating ~ study + C_aspect + C_magnitude + (1|ideophone), data=d)
m5 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + (1|ideophone), data=d)
m6 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + C_length + (1|ideophone), data=d)
m7 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + C_length + C_irregular + (1|ideophone), data=d)
m8 <- lmer(rating ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)

anova(m0,m3) # aspect helps
anova(m3,m4) # magnitude doesn't additionally help
anova(m4,m5) # modality helps
anova(m3,m5) # modality helps also relative to m3
anova(m5,m6) # length helps
anova(m6,m7) # irregular doesn't additionally help
anova(m5,m8) # aspect + modality + length is where it's at
# so m8 is the best performing most complex model
anova(m0,m8)
summary(m8)
fixef(m8)

```

### Control analysis: results stand even when leaving out C_modality

The `C_modality` congruence measure is intended to capture the fact that spoken words can show a range of acoustic similarities to sonic events, independently from other structural correspondences (in other words, it separates imagic iconicity from Gestalt iconicity and relative iconicity). We don't code for specific acoustic correspondences, because these would be by nature only apply to ideophones evocative of sound and would greatly complicate the coding scheme.

This makes it the least fine-grained coding category. And given that it applies to a semantic domain that is known, from prior work and first principles, to be easiest to guess, one could think of `C_modality` as just giving sound words a leg up in the cumulative scores. Therefore, as a control, it is useful to know whether the patterns we find hold up even if we disregard `C_modality` entirely.

```{r sound_off}

# we add C_nosound as a cumulative measure that excludes modality
d <- d %>%
  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude)

# recreate dp to include C_nosound
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

d %>%
  group_by(C_nosound) %>%
  summarise(n=n(),score=mean.na(score))

ggplot(data=dp, aes(x=C_nosound,y=score_z,colour=C_nosound)) +
  theme_tufte() + theme(legend.position="none") +
  labs(title="Guessability (both studies) by cumulative iconicity",
       x="cumulative iconicity (excluding modality)",
       y=expression('guessability ('~italic(z)~')')) +
  geom_half_boxplot() +
  geom_half_point() +
  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
  NULL

```

There is an r=0.26 correlation between logodds and C_sound across the studies. This is present in both studies.

```{r congruency_stats_nosound_1}

# simplest thing we can do is a correlation.

cor.test(d$logodds,d$C_nosound,method="pearson")

d.Collabra <- d %>% filter(study == "Collabra")
d.Language <- d %>% filter(study == "Language")
cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")
cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")

p0 <- cor.test(d$logodds,d$C_nosound,method="pearson")$p.value
p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")$p.value
p2 <- cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")$p.value
p.adjust(c(p0,p1,p2),method="bonferroni")

```

The mixed effects models look much the same.
A null model `m0` predices logodds by paper, controlling for item-level differences. A model `m1` with cumulative congruence as an additional fixed effect fares significantly better.

For the fixed effects, m1 lists a 0.205 improvement in logodds for `C_nosound`. According to my interpretation this means that every added iconic feature leads to an ~8% boost in proportion correct. The lower baseline for the Language set is estimated as -0.513, which translates to a 19.5% guessability penalty.

```{r congruency_stats_nosound2}

m0 <- lmer(logodds ~ study + (1|ideophone), data=d)
m1 <- lmer(logodds ~ study + C_nosound + (1|ideophone), data=d)
anova(m0,m1)

fixef(m1)

probitlink(0.205,inverse=T)-0.5
probitlink(-0.513,inverse=T)-0.5

```

## Exploring coding sensitivity and congruency blind spots
What are we not capturing? Our coding and congruency measures are designed to be reproducible and human-codable. They are not exhaustive, and there may well be aspects that we are missing. Some ideophones may also be guessed relatively high for reasons other than form-meaning analogies (e.g., they had a very unattractive foil or presented unforeseen confounds). Here we explore the congruency data to find highly guessed ideophones that our coding did not (yet) identify as having iconic mappings. We also look at the other end: ideophones coded as having an iconic mapping but not doing well in the binary forced choice task.

If we sort by score most of the blindspots are from the Collabra study, which had a less forbidding design and therefore a higher baseline score.

```{r congruency_blindspots}

d %>%
  filter(C_cumulative == 0, score_z > 0.5) %>%
  dplyr::select(filename,ideophone,meaning,meaning_NL,language,logodds,score) %>%
  arrange(-score) %>% ungroup %>%
  slice(1:10)

```

Coding sensitivity varies by feature, in part because not all are equally frequent. For instance, `C_weight_tone` captures ideophones that have high tone/intonation and small/light meaning or conversely low tone/intonation and big/heavy meaning.

```{r coding_sensitivity}

icoplot("C_weight_tone")

d %>%
  group_by(language) %>%
  filter(C_weight_tone == 1) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))

d %>%
  group_by(language) %>%
  filter(C_weight_tone == 1) %>%
  View()

```

### Variation by language

Although we have focused on cross-linguistically attested features, features may also vary by language. For instance, irregularity may work somewhat in Japanese, Korean and Ewe, the 4 Semai ideophones scored as showing this mapping have a mean accuracy score of 0.45, dragging down the average. It looks like partial reduplication really has a different function in Semai, and one that is not captured by our coding scheme.

```{r coding_by_language}

d %>%
  group_by(language) %>%
  filter(C_irregular == 1) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating),C=mean.na(C_cumulative))

d %>%
  group_by(language) %>%
  filter(C_irregular == 1) %>%
  dplyr::select(ideophone,meaning,language,category,score,rating,C_irregular,C_cumulative) %>%
  View()

mS0 <- lmer(logodds ~ study + (1|ideophone), data=d)
mS1 <- lmer(logodds ~ study + language + (1|ideophone), data=d) # fails to converge
mS2 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)
mS3 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + language + (1|ideophone), data=d)

anova(mS2,mS3)

fixef(mS3)

```

One might conclude from that language would be a relevant factor to take into account in the statistical models. But its influence is not very large. If we add it to the most complete model above, the fit is marginally improved, but the difference is not of the same magnitude as other factors. Also, language is partly confounded with quality of recordings, which are lab recordings for Ewe, Korean, Japanese and field recordings for Siwu and Semai. We can dummy code this as a new variable to check its effect.

```{r coding_by_audio}

d <- d %>%
  mutate(audio = ifelse(language %in% c("Siwu","Semai"),"field","lab"))

d %>%
  group_by(audio) %>%
  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))

ggplot(data=d, aes(x=audio,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by audio version",
       x="audio",
       y="guessability (% correct)") +
  geom_beeswarm(cex=2.5) +
  NULL

ggplot(data=d, aes(x=language,y=score,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
  labs(title="Guessability by audio version",
       x="language",
       y="guessability (% correct)") +
  geom_beeswarm(cex=2.5) +
  NULL

# ratings look like they have a distinctly non-normal distribution for field recordings, with a large bulge below 3 (= middle)
ggplot(data=d, aes(x=audio,y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       y="mean iconicity rating") +
  geom_beeswarm(cex=2.5) +
  NULL

# and indeed the bulge isn't really there for any of the non-field languages (and arguably mostly attributable to the Semai recordings)
ggplot(data=d, aes(x=language,y=rating,fill=C_cumulative,colour=C_cumulative)) +
  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
  labs(title="Iconicity ratings for 239 ideophones",
       x="language",
       y="mean iconicity rating") +
  geom_beeswarm(cex=2.5) +
  NULL

```

```{r congruency_underperformers}
# the reverse is also interesting: which ideophones were coded as having iconic mappings yet were not guessed greatly?

d %>%
  filter(C_cumulative > 0, logodds < -0.25) %>%
  dplyr::select(filename,ideophone,meaning,meaning_NL,C_cumulative,language,logodds,score) %>%
  arrange(score) %>% ungroup() %>%
  slice(1:10)

```

### 2021-06-09 TI_functions.R

@@ -18,7 +18,7 @@ icoplot <- function(var=NULL,saveplot=FALSE) {
     theme_tufte() + ylim(0,1) + 
     theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0))
   # set colour scheme based on nature of variable
-  if (var %in% c("C_cumulative","C_ternary","language","C_simple")) {
+  if (var %in% c("C_cumulative","C_ternary","language","C_simple","C_nosound")) {
     p <- p +
       scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
       scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9)

### 2022-01-17 TI_03_ratings_BM.Rmd
this runs now but Im not sure what the graph in the chunk i relabelled problemchunk is meant to be

--- 2021-06-09 TI_03_ratings.Rmd
@@
                       echo=TRUE, warning=FALSE, message=FALSE)
-
 ```
@@
 # Packages
-list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","viridis","gghalves","plotly","lme4","VGAM","plyr","emmeans","car")
-new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
-if(length(new.packages)) install.packages(new.packages)
-lapply(list.of.packages, require, character.only=T)
+library(tidyverse)
+library(readxl)
+library(plyr)
+library(ggthemes)
+library(gghalves)
+library(car)
+library(emmeans)
+library(plotly)
+library(ggpubr)
+library(viridis)
 
@@
 # check for inconsistent raters with person-total correlation based on Motamedi et al. (2019)
+
 ptc <- d.uncorr[,c("item","rater","rating")]
@@
   xlab("rating") + ylab("guessability")
-ggsave("figures\\rating_guessing_corr.png",height=5,width=7.5)
+#ggsave("figures\\rating_guessing_corr.png",height=5,width=7.5)
 
@@
 
-```{r}
+```{r problem chunk}
 
@@
 
-plot_ly(data=d,x=rating_z,y=score_z,color=C_simple)
+#plot_ly(data=d,x=rating_z,y=score_z,color=C_simple)
 

### 2022-01-31 paper_coding_2_analysis.Rmd
started making shiny app to explore data

@@
 # write this dataset as the fullest dataset
-write.csv(d,file="data/ideophones_coded_guessed_rated.csv")
+write.csv(d,file="data/ideophones_coded_guessed_rated.csv",fileEncoding = "UTF-8")
 write_xlsx(d,path="data/ideophones_coded_guessed_rated.xlsx")

### 2022-01-31 visualisations_bonnie.R

@@ -1,0 +1,100 @@
+library(tidyverse)
+library(plotly)
+library(ggthemes)
+library(shiny)
+library(readxl)
+library(VGAM)
+library(viridis)
+
+#### For some reason writing the d dataset from paper_coding_2_analysis.Rmd
+#### to utf-8 and then reading it in here as a utf-8 file didn't work :(
+
+#### So I just make it again here because that works
+# get consensus coding data
+d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)
+
+# add guessability scores from the Collabra and Language studies.
+
+d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
+  dplyr::select(-category)
+d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))
+
+# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
+d <- d %>%
+  group_by(study) %>%
+  mutate(logodds = probitlink(score))
+
+# add Z score to make scores more comparable in plots across studies
+d <- d %>%
+  group_by(study) %>%
+  mutate(score_z = scale(score,center=T,scale=T))
+
+# get ratings data
+d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
+  dplyr::select(-category,-list,-item)
+d <- left_join(d,d.ratings,by=c("filename","study","language"))
+
+# add z score for ratings
+d$rating_z <- scale(d$rating,center=T,scale=T)
+
+d%>%
+  mutate(label=paste(ideophone,paste("'",meaning,"'",sep=""),language))->dat
+  
+# Define UI for app
+ui <- fluidPage(
+  titlePanel("Triangulating iconicity"),
+  sidebarLayout(
+    
+    # Sidebar panel for inputs ----
+    sidebarPanel(
+      selectInput("semdom","Choose a semantic domain",c("All domains",dat$category)),
+      selectInput("lang","Choose a language",c("All languages",dat$language)),
+      selectInput("study","Choose a study",c("Both studies",dat$study)),
+      checkboxInput("pointlabels","Label points",FALSE)),
+      
+    
+    # Main panel for displaying outputs ----
+    mainPanel(
+      
+      plotlyOutput(outputId = "plot")
+      
+    )
+  )
+)
+
+server <- function(input, output) {
+  
+  output$plot <- renderPlotly({
+    
[…]

### 2022-01-31 visualisations_bonnie.R
adding more informative labels

@@ -38,11 +38,17 @@ d <- left_join(d,d.ratings,by=c("filename","study","language"))
 d$rating_z <- scale(d$rating,center=T,scale=T)
 
 d%>%
-  mutate(label=paste(ideophone,paste("'",meaning,"'",sep=""),language))->dat
-  
+  pivot_longer(C_modality:C_magnitude,names_to="correlate",values_to="present")%>%
+  filter(present!=0)%>%
+  mutate(features="features")%>%
+  group_by(ideophone)%>%
+  pivot_wider(names_from="features",values_from="correlate",values_fn=list)%>%
+  mutate(label=paste(ideophone,meaning,features,sep="\n"))->dat
+
 # Define UI for app
 ui <- fluidPage(
   titlePanel("Triangulating iconicity"),
+  fluidRow(includeHTML("test.html")),
   sidebarLayout(
     
     # Sidebar panel for inputs ----
@@ -79,7 +85,6 @@ server <- function(input, output) {
       filter(language %in% langs)%>%
       filter(category %in% semdoms)->filtered_dat
     
-    print(input$labelpoints)
     if(input$pointlabels==TRUE){
     filtered_dat%>%
       ggplot(aes(x=rating_z,y=logodds,colour=C_cumulative,label=label))+geom_point()+scale_fill_viridis(option="plasma") +

### 2022-01-31 visualisations_bonnie.R
adding more informative labels

@@ -48,7 +48,7 @@ d%>%
 # Define UI for app
 ui <- fluidPage(
   titlePanel("Triangulating iconicity"),
-  fluidRow(includeHTML("test.html")),
+  fluidRow(includeHTML("codingscheme.html")),
   sidebarLayout(
     
     # Sidebar panel for inputs ----

### 2022-01-31 visualisations_bonnie.R
removing features not used

@@ -38,7 +38,8 @@ d <- left_join(d,d.ratings,by=c("filename","study","language"))
 d$rating_z <- scale(d$rating,center=T,scale=T)
 
 d%>%
-  pivot_longer(C_modality:C_magnitude,names_to="correlate",values_to="present")%>%
+  select(-C_length,-C_aspect,-C_magnitude,-C_weight)%>%
+  pivot_longer(C_modality:C_weight_tone,names_to="correlate",values_to="present")%>%
   filter(present!=0)%>%
   mutate(features="features")%>%
   group_by(ideophone)%>%

### 2024-01-16 TI_03_ratings.Rmd
update & spring cleaning

@@
 title: "Triangulating iconicity: Iconicity ratings"
-author: "Mark Dingemanse & Stella Punselie"
-date: "Updated `r format(Sys.Date())`"
+author: "Stella Punselie & Mark Dingemanse"
+date: "(updated `r format(Sys.Date())`)"
 output: github_document
@@
 }
-View(personTotalCorrelationCorrected(ptc))
+personTotalCorrelationCorrected(ptc)
 
@@
   "Texture","Colour/Visual"))
-ggsave("figures\\rating_category.png",height=5,width=7.5)
+#ggsave("figures\\rating_category.png",height=5,width=7.5)
 
@@
   xlab("category")
-ggsave("figures\\rating_language.png",height=5,width=7.5)
+#ggsave("figures\\rating_language.png",height=5,width=7.5)
 

### 2024-01-16 paper_coding_2_analysis.Rmd
update & spring cleaning

--- 2021-06-09 paper_coding_2_analysis.Rmd
@@
 
+# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
+tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
+dp <- d
+dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
+
 # blank plot can be useful as a canvas 
@@
   theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
-  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
+  stat_summary(fun=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
   geom_dotplot(colour="white",fill="black",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
@@
 ```{r visuals}
-
-# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
 

### 2024-04-03 paper_01_consistency.Rmd
knitted md for github

--- 2021-06-09 paper_coding_1_consistency.Rmd
@@
 
-d1 = read_excel("data\\ideophones_coded_MD_original.xlsx") %>% arrange(filename)
-d2 = read_excel("data\\ideophones_coded_SP_original.xlsx") %>% arrange(filename)
+d1 = read_excel("data/ideophones_coded_MD_original.xlsx") %>% arrange(filename)
+d2 = read_excel("data/ideophones_coded_SP_original.xlsx") %>% arrange(filename)
 d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge
@@
 F_consistency$kappa <- NA
-for (i in 1:(length(F_categories)-1)) { # -1 to exclude the notes column
+for (i in 1:(length(F_categories)-1)) { 
   F_consistency$agree[i] <- (get_agree(data=d.F,cat=F_categories[i]))
@@
 M_consistency
+mean.na(M_consistency$agree)
 
@@
 
-The original coding sheet is available as X and the updated coding sheet as X_version_2.
+The original coding sheet is available as `predico_coding_scheme_1.pdf` and the updated coding sheet as `predico_coding_scheme_2.pdf`.
 
@@
 ```{r consistency_2ndround}
-d1 = read_excel("data\\ideophones_coded_MD.xlsx") %>% arrange(filename)
-d2 = read_excel("data\\ideophones_coded_SP.xlsx") %>% arrange(filename)
+d1 = read_excel("data/ideophones_coded_MD.xlsx") %>% arrange(filename)
+d2 = read_excel("data/ideophones_coded_SP.xlsx") %>% arrange(filename)
 d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge
@@
 
-write_xlsx(d.F.inconsistent,path="data\\ideophones_F_consistency_2ndround.xlsx")
+write_xlsx(d.F.inconsistent,path="data/ideophones_F_consistency_2ndround.xlsx")
 
@@
 
-d.coded <- read_excel("data\\ideophones_coded_SP.xlsx") %>% 
+d.coded <- read_excel("data/ideophones_coded_SP.xlsx") %>% 
   arrange(filename) %>%
@@
 
-# we can have a first look at the basic distribution of guessability by iconicity
-d %>%
-  group_by(C_cumulative) %>%
-  summarise(n=n(),score=mean.na(score))
+```
 
-d %>%
-  group_by(C_any1) %>%
-  summarise(n=n(),score=mean.na(score))
+Finally, we write this to XLSX and CSV files. (Commented out here to avoid overwriting anytime we knit the markdown.)
 
-d %>%
-  group_by(C_atleast2) %>%
-  summarise(n=n(),score=mean.na(score))
+```{r write_files}
 
-d %>%
-  group_by(C_ternary) %>%
-  summarise(n=n(),score=mean.na(score))
-
-d %>%
-  group_by(C_simple) %>%
-  summarise(n=n(),score=mean.na(score))
-
-# finally write this to XLS and CSV
-write_xlsx(d,path="data\\ideophones_coded.xlsx")
-write.csv(d,file="data\\ideophones_coded.csv")
+#write_xlsx(d,path="data\\ideophones_coded.xlsx")
+#write.csv(d,file="data\\ideophones_coded.csv")
 

### 2024-04-03 paper_02_main_analyses.Rmd
rename things and clean up

--- 2024-01-16 paper_coding_2_analysis.Rmd
@@
 title: "Triangulating iconicity: Coding analysis"
-author: "Mark Dingemanse & Stella Punselie"
+author: "Mark Dingemanse, Stella Punselie, Bonnie McLean"
 date: "Updated `r format(Sys.Date())`"
@@
 # get consensus coding data
-d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)
+d = read_excel("data/ideophones_coded.xlsx") %>% arrange(filename)
 
@@
 
-d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
+d.scores = read_excel("data/ideophones_guessability.xlsx") %>%
   dplyr::select(-category)
@@
 # get ratings data
-d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
+d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") %>%
   dplyr::select(-category,-list,-item)
-d <- left_join(d,d.ratings,by=c("filename","study","language"))
+d <- left_join(d,d.ratings,by=c("filename","study","language")) |>
+  mutate(filename = gsub("_org","",filename))
 
@@
 
+ideophones_filename <- d |> ungroup() |> select(ideophone,filename) |>
+  mutate(filename = str_to_lower(filename))
+
+# load the full ratings data for some of the reporting in the paper
+d.ratings.full <- read_xlsx("data\\ideophones_rated.xlsx") |>
+  mutate(list = gsub(".*([A-Z])$", "\\1",item)) |>
+  mutate(filename = gsub("^(.*)_.*$", "\\1", str_to_lower(item))) |>
+  mutate(filename = gsub("Coll_","", filename)) |>
+  select(-item)
+
+# match the data to ideophones for easier lookup
+d.ratings.full <- left_join(d.ratings.full,
+                  ideophones_filename,by="filename",copy=TRUE)
+
 ```
@@
   group_by(study) %>%
-  summarize(max=max(score),min=min(score),mean=mean.na(score))
+  dplyr::summarize(max=max(score),min=min(score),mean=mean.na(score))
 
@@
 
-ggsave("figures\\paper-guessability-blank.pdf",height=5,width=7.5)
-
-d %>% filter(filename=="Korean_23_Motion_org") %>% View()
+ggsave("figures\\fig2-guessability-blank.pdf",height=5,width=7.5)
 
@@
 
-### Figure: Guessability by study and iconicity rating
+Let's plot the guessability by study and iconicity rating
 
 ```{r figures_paper_guessability_rating}
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
 
@@
 plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)
+#ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)
 
@@
 
-### Figure: Guessability and cumulative iconicity
+### Figure 3: Guessability and cumulative iconicity
 
 ```{r figures_paper_guessability_coding}
-
-# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
 
@@
 plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\paper_panelAB.png",height=5,width=9)
+ggsave("figures\\fig3-panelAB.png",height=4,width=9,bg="white")
 
@@
 
-### Figure: Ratings and cumulative iconicity
+### Figure 4: Ratings by study (A) and by semantic domain (B)
+
+Panel figure for Study B providing a view of ratings per study and per semantic domain.
+
+```{r fig4_panel}
+
+#panel A: ratings by study
+pA <- d |>  ggplot(aes(x=score_z,y=rating)) +   
+  theme_tufte(base_size = 16) +
+  theme(legend.position = c(0.2,0.9),
+        legend.title = element_blank()) +
+  scale_fill_manual(values=c("white","black")) +
+  geom_point(aes(fill=study),
+             position="jitter",
+             shape=21,
+             colour="black") +
+  geom_smooth(method=loess,colour="black",alpha=0.5) +
+  xlab("guessability (z)") + ylab("rating") +
+  NULL
+
+#panel B: ratings by semantic domain
+pB <- d |>
+  filter(category != "Other") |>
+  ggplot(aes(x=reorder(category,-rating),y=rating,color=category)) +
+  theme_tufte(base_size=16) +
+  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
+  geom_half_point(show.legend=F) +
+  scale_colour_viridis_d(option="D",alpha=0.8) +
+  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
+  "Texture","Colour/Visual"))
+
+plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.6,2))
+ggsave("figures/fig4-panel_ratings.png",height=4,width=9,bg="white")
+
+# a correlation test shows a Pearson correlation of .57 between iconicity rating
+# and guessability
+cor.test(d$rating_z,d$logodds,method="pearson")
+
+```
+
+Table 3: Examples of ideophones from the low/high rating/guessability quadrants
+
+```{r example_ratings_guessables}
+
+# these examples are given in Table 3 in the paper
+examples <- c("ton ton","kpa",
+              "slʔẽẽk","miɔmiɔ",
+              "kodzokodzo","ɟtoonɟtoonɟtoon",
+              "fututu", "plɒ̃s")
+d |>
+  filter(ideophone %in% examples) |>
+  group_by(ideophone,language) |>
+  dplyr::summarise(n=n(),
+                   mean_rating = mean(rating),
+                   score = score)
+
+```
+
+### Figure 5: Ratings and cumulative iconicity
 
@@
 plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\paper_panelAB_ratings.png",height=5,width=9)
-
-```
-
-### Figure: Mappings and semantic domains
-
-```{r figure_mappings_domains}
-
-library(forcats)
-dp.categories <- d %>%
-  filter(category %notin% "Other") %>%
-  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
-  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
-  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
-  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis
-
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-dp.categories %>%
-  ungroup() %>%
-  group_by(category_binary) %>%
-  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
-group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
-t.test(group1,group2)
-library(effsize)
-cohen.d(group1,group2)
-
-# also when we exclude C_modality? yes
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-dp.categories %>%
-  group_by(category_binary) %>%
-  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
-group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
-t.test(group1,group2)
-cohen.d(group1,group2)
-
-ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
-  theme_tufte() + theme(legend.position="none") +
-  labs(title="Cumulative iconicity by semantic domain") +
-  geom_half_violin() +
-  geom_half_point() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  NULL
-
-ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
-  theme_tufte() + theme(legend.position="none") +
-  labs(title="Cumulative iconicity by semantic domain") +
-  geom_hex() +
-  NULL
-
-```
-
-### Raincloud plots
-While the icoplots provide a useful first visualisation, for in a paper we would probably want to use something like raincloud plots that show key features of the distribution across groups.
-
-```{r visuals}
-
-# panel A: scores by study and cumulative iconicity
-ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank()) + 
-  labs(title="Guessability by study and cumulative iconicity") +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
-  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
-
-ggplot(data=dp, aes(x=C_cumulative,y=logodds)) +
-  theme_tufte() +
-  labs(title="Guessability (log odds) by cumulative iconicity",
-       x="cumulative iconicity",
-       y="guessability (log odds)") +
-  geom_half_boxplot() +
-  geom_half_point() +
-  NULL
-
-ggplot(data=dp, aes(x=C_any1,y=logodds)) +
-  theme_tufte() +
-  geom_half_boxplot() +
-  geom_half_point() +
-  NULL
-
-ggplot(data=dp, aes(x=C_atleast2,y=logodds)) +
-  theme_tufte() +
-  geom_half_boxplot() +
-  geom_half_point() +
-  NULL
-
-ggplot(data=dp, aes(x=C_ternary,y=logodds)) +
-  theme_tufte() +
-  geom_half_boxplot() +
-  geom_half_point() +
-  NULL
-
-ggplot(data=dp, aes(x="all",y=score_z)) +
-  theme_tufte() +
-  geom_half_boxplot() +
-  geom_half_point(side="l") +
-  geom_half_violin(side="r") +
-  NULL
-
-ggplot(data=dp, aes(x=C_cumulative,y=logodds,color=C_cumulative,fill=C_cumulative)) +
-  theme_tufte() +
-  geom_half_violin() +
-  geom_half_point_panel() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9)
-
-icoplot("C_cumulative")
-
-ggplot(data=dp, aes(x=C_cumulative,y=logodds,color=C_cumulative,fill=C_cumulative)) +
-  theme_tufte() +
-  geom_half_violin() +
-  geom_half_point_panel() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  facet_grid(~ study)
-
-```
-
-### Dotplots
-
-A quick look at some congruency measures. Uses a custom plotting function `icoplot()`, which is a wrapper for a ggplot object constructed with `geom_dotplot`.
-
-```{r icoplots}
-
-# all data points by language
-icoplot() # icoplot() is loaded from TI_functions.R
-
-icoplot("C_modality")
-icoplot("C_aspect")
-icoplot("C_length")
-icoplot("C_magnitude")
-icoplot("C_irregular")
-icoplot("C_cumulative")
-
-# how do the constituents of C_magnitude fare?
-icoplot("C_magnitude")
-icoplot("C_weight_voice")
-icoplot("C_weight_vowel")
-icoplot("C_weight_tone")
-
-icoplot("C_any1")
-icoplot("C_atleast2")
-icoplot("C_ternary")
-icoplot("C_simple")
-
-# blank plot can be useful as background for slides
-ggplot(data=dp, aes(x=study,y=score)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
-  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
-  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
-
-# facet by language to discover crosslinguistic diversity
-ggplot(data=dp, aes(x=study,y=score)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
-  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
-  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
+ggsave("figures/fig6-panelAB_ratings.png",height=5,width=9,bg="white")
 
@@
 
-The simplest thing to do is a simple correlation. Looks like there is an r=0.30 correlation between logodds and C_cumulative.
+The simplest thing to do is a correlation.
 
@@
 p.adjust(c(p0,p1,p2),method="bonferroni")
+pvalues <-p.adjust(c(p0,p1,p2),method="bonferroni")
 
@@
 
+# fixed effects
 fixef(m1)
 
-probitlink(0.1823,inverse=T)-0.5
-probitlink(-0.5242,inverse=T)-0.5
+# turn this into a percentage under the simplifying assumption that the distance
+# between iconicity measures is the same at every scale point
+probitlink(fixef(m1)[[3]],inverse=T)-0.5
+probitlink(fixef(m1)[[2]],inverse=T)-0.5
 
@@
 probitlink(fixef(m8)[2],inverse=T)-0.5
-probitlink(fixef(m8)[3],inverse=T)-0.5
-probitlink(fixef(m8)[4],inverse=T)-0.5
-probitlink(fixef(m8)[5],inverse=T)-0.5
+probitlink(fixef(m8)[3],inverse=T)-0.5 # aspect
+probitlink(fixef(m8)[4],inverse=T)-0.5 # modality
+probitlink(fixef(m8)[5],inverse=T)-0.5 # length
 
@@
 
-### Semantic domains
+### Structural correspondences and semantic domains
+
 Our prediction is that Sound and Motion ideophones should show more structural correspondences than the other domains. This is borne out.
 
-```{r coding_semanticdomains}
+```{r descriptive_semantic_domains}
 
-library(forcats)
 dp.categories <- d %>%
   filter(category %notin% "Other") %>%
-  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
+  mutate(category = forcats::fct_reorder(category, -C_cumulative, .fun='mean')) %>%
   mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
@@
 
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
+# mean of C_cumuluative
+dp.categories |>
+  group_by(category) |>
+  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
   arrange(-mean_C)
 
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
+# mean of only the strongest predictors
+dp.categories |>
+  group_by(category) |>
+  dplyr::summarise(n=n(),mean_strongest=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
+  arrange(-mean_strongest)
 
-dp.categories %>%
-  ungroup() %>%
+# binary comparison
+dp.categories |>
+  ungroup() |>
   group_by(category_binary) %>%
-  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
+  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
   arrange(-mean_C)
@@
 t.test(group1,group2)
-library(effsize)
-cohen.d(group1,group2)
+effsize::cohen.d(group1,group2)
 
 # also when we exclude C_modality? yes
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
+dp.categories |>
+  group_by(category) |>
+  dplyr::summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
   arrange(-mean_C)
-dp.categories %>%
-  group_by(category_binary) %>%
-  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
+dp.categories |>
+  group_by(category_binary) |>
+  dplyr::summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
   arrange(-mean_C)
@@
 t.test(group1,group2)
-cohen.d(group1,group2)
+effsize::cohen.d(group1,group2)
 
@@
 
-We can also look in more detail at the relative differences by semantic domain by building a mixed model that predicts cumulative iconicity with domain as a predictor.
+### Study B: Rating iconicity
 
-```{r coding_category_lme}
+Here we look at the iconicity ratings for all ideophones across both studies.
 
-m0 <- lm(C_cumulative ~ category, data=d)
-effsizes(m0)
+```{r coding_ratings_descriptives}
+
+d.ratings.full |>
+  group_by(ideophone,language) |>
+  dplyr::summarise(n=n(),mean=mean.na(rating),sd=sd.na(rating))
+
+# get means and standard deviations across all ideophones and for each language and category
+mean(d.allratings$rating)
+  # 2.948339
+sd(d.allratings$rating)
+
+examples <- c("kpa","boo boo","zuratto", "plɒ̃s")
+d.ratings.full |>
+  filter(ideophone %in% examples) |>
+  group_by(ideophone,language) |>
+  dplyr::summarise(n=n(),
+                   mean_rating = mean(rating),
+                   sd_rating = sd(rating))
 
@@
 
-Another prediction is that Sound and Motion ideophones should receive higher iconicity ratings than other domains. This prediction is also borne out.
+There appears to be a positive correlation between the rating and guessing data.
 
-```{r rating_domains}
+### Ratings: numbers
 
-group1 <- subset(dp.categories, category_binary == "SoundMotion")$rating
-group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$rating
-t.test(group1,group2)
-cohen.d(group1,group2)
+Iconicity ratings by semantic domain
+
+```{r ratings_by_domain}
+
+d.ratings.full |>
+  filter(category != "Other") |>
+  group_by(category) |>
+  dplyr::summarise(n=n(),
+                   mean_rating=mean(rating),
+                   sd_rating = sd(rating)) |>
+  arrange(desc(mean_rating))
 
@@
 
-### Coding and ratings
-
-As above, the simplest thing to do is run a simple correlation: there is an r=0.43 correlation between meaning iconicity ratings and cumulative iconicity.
-
-```{r coding_ratings}
-
-cor.test(d$rating_z,d$C_cumulative,method="pearson")
+```{r linear_regressions}
 
@@
 
+How do iconicity ratings relate to structural correspondences?
+
+A simple correlation test suggests a positive relation between iconicity rating and cumulative iconicity.
 As above, we can go a bit deeper by using mixed effects modelling.
 We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away.
-However, the studies are also different in terms of
 
@@
 
+# simple correlation test
+cor.test(d$rating_z,d$C_cumulative,method="pearson")
+
+# lme
 m0 <- lmer(rating ~ study + (1|ideophone), data=d)
@@
 
-We can assess the relative weight of different features using model comparison.
+As above, we can assess the relative weight of different features using model comparison.
 We build models by adding the congruency measures in rank order of attestation in the coded data.
@@
 ```
-
-### Control analysis: results stand even when leaving out C_modality
-
-The `C_modality` congruence measure is intended to capture the fact that spoken words can show a range of acoustic similarities to sonic events, independently from other structural correspondences (in other words, it separates imagic iconicity from Gestalt iconicity and relative iconicity).
-We don't code for specific acoustic correspondences, because these would be by nature only apply to ideophones evocative of sound and would greatly complicate the coding scheme.
-
-This makes it the least fine-grained coding category.
-And given that it applies to a semantic domain that is known, from prior work and first principles, to be easiest to guess, one could think of `C_modality` as just giving sound words a leg up in the cumulative scores.
-Therefore, as a control, it is useful to know whether the patterns we find hold up even if we disregard `C_modality` entirely.
-
-```{r sound_off}
-
-# we add C_nosound as a cumulative measure that excludes modality
-d <- d %>%
-  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude)
-
-# recreate dp to include C_nosound
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
-
-d %>%
-  group_by(C_nosound) %>%
-  summarise(n=n(),score=mean.na(score))
-
-ggplot(data=dp, aes(x=C_nosound,y=score_z,colour=C_nosound)) +
-  theme_tufte() + theme(legend.position="none") +
-  labs(title="Guessability (both studies) by cumulative iconicity",
-       x="cumulative iconicity (excluding modality)",
-       y=expression('guessability ('~italic(z)~')')) +
-  geom_half_boxplot() +
-  geom_half_point() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  NULL
-
-```
-
-There is an r=0.26 correlation between logodds and C_sound across the studies. This is present in both studies.
-
-```{r congruency_stats_nosound_1}
-
-# simplest thing we can do is a correlation.
-
-cor.test(d$logodds,d$C_nosound,method="pearson")
-
-d.Collabra <- d %>% filter(study == "Collabra")
-d.Language <- d %>% filter(study == "Language")
-cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")
-cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")
-
-p0 <- cor.test(d$logodds,d$C_nosound,method="pearson")$p.value
-p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")$p.value
-p2 <- cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")$p.value
-p.adjust(c(p0,p1,p2),method="bonferroni")
-
-```
-
-The mixed effects models look much the same.
-A null model `m0` predices logodds by paper, controlling for item-level differences. A model `m1` with cumulative congruence as an additional fixed effect fares significantly better.
-
-For the fixed effects, m1 lists a 0.205 improvement in logodds for `C_nosound`.
-According to my interpretation this means that every added iconic feature leads to an ~8% boost in proportion correct.
-The lower baseline for the Language set is estimated as -0.513, which translates to a 19.5% guessability penalty.
-
-```{r congruency_stats_nosound2}
-
-m0 <- lmer(logodds ~ study + (1|ideophone), data=d)
-m1 <- lmer(logodds ~ study + C_nosound + (1|ideophone), data=d)
-anova(m0,m1)
-
-fixef(m1)
-
-probitlink(0.205,inverse=T)-0.5
-probitlink(-0.513,inverse=T)-0.5
-
-```
-
-## Exploring coding sensitivity and congruency blind spots
-What are we not capturing?
-Our coding and congruency measures are designed to be reproducible and human-codable.
-They are not exhaustive, and there may well be aspects that we are missing.
-Some ideophones may also be guessed relatively high for reasons other than form-meaning analogies (e.g., they had a very unattractive foil or presented unforeseen confounds).
-Here we explore the congruency data to find highly guessed ideophones that our coding did not (yet) identify as having iconic mappings.
-We also look at the other end: ideophones coded as having an iconic mapping but not doing well in the binary forced choice task.
-
-If we sort by score most of the blindspots are from the Collabra study, which had a less forbidding design and therefore a higher baseline score.
-
-```{r congruency_blindspots}
-
-d %>%
-  filter(C_cumulative == 0, score_z > 0.5) %>%
-  dplyr::select(filename,ideophone,meaning,meaning_NL,language,logodds,score) %>%
-  arrange(-score) %>% ungroup %>%
-  slice(1:10)
-
-```
-
-Coding sensitivity varies by feature, in part because not all are equally frequent.
-For instance, `C_weight_tone` captures ideophones that have high tone/intonation and small/light meaning or conversely low tone/intonation and big/heavy meaning.
-
-```{r coding_sensitivity}
-
-icoplot("C_weight_tone")
-
-d %>%
-  group_by(language) %>%
-  filter(C_weight_tone == 1) %>%
-  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))
-
-d %>%
-  group_by(language) %>%
-  filter(C_weight_tone == 1) %>%
-  View()
-
-```
-
-### Variation by language
-
-Although we have focused on cross-linguistically attested features, features may also vary by language.
-For instance, irregularity may work somewhat in Japanese, Korean and Ewe, the 4 Semai ideophones scored as showing this mapping have a mean accuracy score of 0.45, dragging down the average.
-It looks like partial reduplication really has a different function in Semai, and one that is not captured by our coding scheme.
-
-```{r coding_by_language}
-
-d %>%
-  group_by(language) %>%
-  filter(C_irregular == 1) %>%
-  summarise(n=n(),score=mean.na(score),rating=mean.na(rating),C=mean.na(C_cumulative))
-
-d %>%
-  group_by(language) %>%
-  filter(C_irregular == 1) %>%
-  dplyr::select(ideophone,meaning,language,category,score,rating,C_irregular,C_cumulative) %>%
-  View()
-
-mS0 <- lmer(logodds ~ study + (1|ideophone), data=d)
-mS1 <- lmer(logodds ~ study + language + (1|ideophone), data=d) # fails to converge
-mS2 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)
-mS3 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + language + (1|ideophone), data=d)
-
-anova(mS2,mS3)
-
-fixef(mS3)
-
-```
-
-One might conclude from that language would be a relevant factor to take into account in the statistical models.
-But its influence is not very large.
-If we add it to the most complete model above, the fit is marginally improved, but the difference is not of the same magnitude as other factors.
-Also, language is partly confounded with quality of recordings, which are lab recordings for Ewe, Korean, Japanese and field recordings for Siwu and Semai.
-We can dummy code this as a new variable to check its effect.
-
-```{r coding_by_audio}
-
-d <- d %>%
-  mutate(audio = ifelse(language %in% c("Siwu","Semai"),"field","lab"))
-
-d %>%
-  group_by(audio) %>%
-  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))
-
-ggplot(data=d, aes(x=audio,y=score,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
-  labs(title="Guessability by audio version",
-       x="audio",
-       y="guessability (% correct)") +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-ggplot(data=d, aes(x=language,y=score,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
-  labs(title="Guessability by audio version",
-       x="language",
-       y="guessability (% correct)") +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-# ratings look like they have a distinctly non-normal distribution for field recordings, with a large bulge below 3 (= middle)
-ggplot(data=d, aes(x=audio,y=rating,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
-  labs(title="Iconicity ratings for 239 ideophones",
-       y="mean iconicity rating") +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-# and indeed the bulge isn't really there for any of the non-field languages (and arguably mostly attributable to the Semai recordings)
-ggplot(data=d, aes(x=language,y=rating,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
-  labs(title="Iconicity ratings for 239 ideophones",
-       x="language",
-       y="mean iconicity rating") +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-```
-
-```{r congruency_underperformers}
-# the reverse is also interesting: which ideophones were coded as having iconic mappings yet were not guessed greatly?
-
-d %>%
-  filter(C_cumulative > 0, logodds < -0.25) %>%
-  dplyr::select(filename,ideophone,meaning,meaning_NL,C_cumulative,language,logodds,score) %>%
-  arrange(score) %>% ungroup() %>%
-  slice(1:10)
-
-```

### 2024-04-03 paper_02_main_analyses.Rmd
cleanup and generate github md for main analyses

@@
 
-# custom functions
-source("TI_functions.R")
 ```
@@
 # get consensus coding data
-d = read_excel("data/ideophones_coded.xlsx") %>% arrange(filename)
+d = read_excel("data/ideophones_coded.xlsx") |> arrange(filename)
 
@@
 
-d.scores = read_excel("data/ideophones_guessability.xlsx") %>%
+d.scores = read_excel("data/ideophones_guessability.xlsx") |>
   dplyr::select(-category)
@@
 # add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
-d <- d %>%
-  group_by(study) %>%
+d <- d |>
+  group_by(study) |>
   mutate(logodds = probitlink(score))
@@
 # add Z score to make scores more comparable in plots across studies
-d <- d %>%
-  group_by(study) %>%
+d <- d |>
+  group_by(study) |>
   mutate(score_z = scale(score,center=T,scale=T))
@@
 # get ratings data
-d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") %>%
+d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") |>
   dplyr::select(-category,-list,-item)
@@
 
-# also write a simplified copy for easy inspection in Excel (helpful when writing)
-d %>%
-  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,rating,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
-  write_xlsx(path="data/ideophones_coded_guessed_rated_simplified.xlsx")
-
+# load and merge the full ratings data for some of the reporting in the paper
 ideophones_filename <- d |> ungroup() |> select(ideophone,filename) |>
@@
 
-# load the full ratings data for some of the reporting in the paper
-d.ratings.full <- read_xlsx("data\\ideophones_rated.xlsx") |>
+d.ratings.full <- read_xlsx("data/ideophones_rated.xlsx") |>
   mutate(list = gsub(".*([A-Z])$", "\\1",item)) |>
@@
 ## Figures
-Here we generate some descriptive stats and create the main figures for the paper.
+Here we generate some descriptive stats and create the main figures for the paper. Note that Figures 1 and 2 in the paper are illustrative and their final versions are not generated in R.
 
@@
 
-d %>%
-  group_by(study) %>%
+d |>
+  group_by(study) |>
   dplyr::summarize(max=max(score),min=min(score),mean=mean.na(score))
@@
 
-### Figure: Overview panel
-The overview panel is created separately but we use the data from the guessability studies as a canvas.
+### Figure 2: Overview panel
+The overview of the triangulation method is created separately in CorelDraw but we use the data from the guessability studies as a canvas.
 
@@
 
-# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
+# make a copy dp of the data in which we convert some measures to factors for easy plotting & faceting
 tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
@@
 
-ggsave("figures\\fig2-guessability-blank.pdf",height=5,width=7.5)
-
-```
-
-Let's plot the guessability by study and iconicity rating
-
-```{r figures_paper_guessability_rating}
-
-# panel A: scores by study and cumulative iconicity
-pA <- ggplot(data=dp, aes(x=study,y=score,fill=study,colour=study)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
-  labs(title="Guessability by study",
-       x="study",
-       y="guessability (% correct)") +
-  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
-  NULL
-
-pB <- ggplot(data=dp, aes(x=rating,y=score,fill=study,colour=study)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") +
-  labs(title="Guessability (both studies) by iconicity rating",
-       x="iconicity rating",
-       y="") +
-  geom_point(size=2.0,shape=21) +
-  geom_smooth(method="lm") +
-  NULL
-
-plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-#ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)
+ggsave("figures/fig2-guessability-blank.pdf",height=5,width=7.5)
 
@@
 plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\fig3-panelAB.png",height=4,width=9,bg="white")
+ggsave("figures/fig3-panelAB.png",height=4,width=9,bg="white")
 
@@
 
-# simplest thing we can do is a correlation.
+# Pearson correlation for both of the studies
+d.Collabra <- d |> filter(study == "Collabra")
+d.Language <- d |> filter(study == "Language")
+correlation.Collabra <- cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")
+correlation.Language <- cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")
 
-cor.test(d$logodds,d$C_cumulative,method="pearson")
-
-d.Collabra <- d %>% filter(study == "Collabra")
-d.Language <- d %>% filter(study == "Language")
-cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")
-cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")
-
-p0 <- cor.test(d$logodds,d$C_cumulative,method="pearson")$p.value
-p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_cumulative,method="pearson")$p.value
-p2 <- cor.test(d.Language$logodds,d.Language$C_cumulative,method="pearson")$p.value
-p.adjust(c(p0,p1,p2),method="bonferroni")
-pvalues <-p.adjust(c(p0,p1,p2),method="bonferroni")
+# Bonferroni-correcting the p values for multiple comparisons
+p1 <- correlation.Collabra$p.value
+p2 <- correlation.Language$p.value
+pvalues <-p.adjust(c(p1,p2),method="bonferroni")
 
 ```
+
+The correlation is found in the Collabra study (Pearson’s r = `r round(correlation.Collabra$estimate,2)`, p = `r pvalues[1]`) as well as in the Language study (Pearson’s r = `r round(correlation.Language$estimate,2)`, p `r pvalues[2]`, p-values Bonferroni-corrected for multiple comparisons).
 
@@
 
-dp.categories <- d %>%
-  filter(category %notin% "Other") %>%
-  mutate(category = forcats::fct_reorder(category, -C_cumulative, .fun='mean')) %>%
-  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
-  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
+dp.categories <- d |>
+  filter(category %notin% "Other") |>
+  mutate(category = forcats::fct_reorder(category, -C_cumulative, .fun='mean')) |>
+  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) |>
+  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) |> # for control analysis
   mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis
@@
   group_by(category) |>
-  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
+  dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
   arrange(-mean_C)
@@
   ungroup() |>
-  group_by(category_binary) %>%
+  group_by(category_binary) |>
   dplyr::summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) |>
@@
 # get means and standard deviations across all ideophones and for each language and category
-mean(d.allratings$rating)
-  # 2.948339
-sd(d.allratings$rating)
+mean(d.ratings.full$rating)
+
+sd(d.ratings.full$rating)
 
@@
 
-```{r linear_regressions}
+How do iconicity ratings relate to structural correspondences?
 
+```{r coding_ratings_correlation}
+correlation_coding_ratings <- cor.test(d$rating_z,d$C_cumulative,method="pearson")
 ```
 
-How do iconicity ratings relate to structural correspondences?
+A simple correlation test suggests a positive relation between iconicity rating and cumulative iconicity: Pearson's *r* = `r round(correlation_coding_ratings$estimate,2)`, p = `r correlation_coding_ratings$p.value`.
 
-A simple correlation test suggests a positive relation between iconicity rating and cumulative iconicity.
-As above, we can go a bit deeper by using mixed effects modelling.
-We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away.
+As above, we can go a bit deeper by using mixed effects modelling. We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away.
 
 ```{r coding_ratings_lme}
-
-# simple correlation test
-cor.test(d$rating_z,d$C_cumulative,method="pearson")
 
@@
 
+m0m1.chisquare <- round(anova(m0,m1)$Chisq[2],2)
+m0m1.pvalue <- anova(m0,m1)$"Pr(>Chisq)"[2]
+m0m1.loglikdiff <- abs(anova(m0,m1)$logLik[1] - anova(m0,m1)$logLik[2])
+
 fixef(m1)
@@
 
-As above, we can assess the relative weight of different features using model comparison.
-We build models by adding the congruency measures in rank order of attestation in the coded data.
-Doing this we find that C_aspect, C_modality and C_length are responsible for the bulk of the variance.
-Compared to the null model that only represents the baseline difference between studies, model `m8` fares significantly better (log likelihood difference = 21.3).
+We find that a model with cumulative iconicity as an additional fixed effect achieves a fit that is significantly better than a null model without it (χ2(1) = `r m0m1.chisquare`, p = `r m0m1.pvalue`, log likelihood difference `r m0m1.loglikdiff`).
+The fixed effect of cumulative iconicity is estimated at `r fixef(m1)[["C_cumulative"]]`.
 
-The model estimates given in the output of `fixef(m8)` represent the contribution of a given fixed effect while controlling for others, but they are NOT additive (i.e.
-can't be summed to arrive at the effect of `C_cumulative`).
-
-* `C_aspect`: 0.181 logodds (+7% boost)
-* `C_modality`: 0.177 logodds (+7% boost)
-* `C_length`: 0.443 logodds (+17% boost)
-
-```{r congruency_stats_3}
+```{r ratings_stats_3}
 
@@
 anova(m0,m8)
-summary(m8)
+
+m0m8.chisquare <- round(anova(m0,m8)$Chisq[2],2)
+m0m8.pvalue <- anova(m0,m8)$"Pr(>Chisq)"[2]
+m0m8.loglikdiff <- abs(anova(m0,m8)$logLik[1] - anova(m0,m8)$logLik[2])
+
+m8.estimates <- summary(m8)[["coefficients"]][,1]
+
 fixef(m8)
@@
 ```
+
+We can also assess the relative weight of different features using model comparison.
+We build models by adding the congruency measures in rank order of attestation in the coded data.
+Doing this we find that C_aspect, C_modality and C_length are responsible for the bulk of the variance.
+Compared to the null model that only represents the baseline difference between studies, model `m8` fares significantly better (Chisquare = `r m0m8.chisquare`, p = `r m0m8.pvalue`, log likelihood difference = `r m0m8.loglikdiff`).
+
+The model estimates given in the output of `fixef(m8)` represent the contribution of a given fixed effect while controlling for others, but they are NOT additive (i.e.
+they cannot be summed to arrive at the effect of `C_cumulative`).
+
+* `C_length`: `r round(m8.estimates["C_length"],2)`
+* `C_modality`: `r round(m8.estimates["C_modality"],2)`
+* `C_aspect`: `r round(m8.estimates["C_aspect"],2)`

### 2024-04-03 legacy/TI_02_congruence.Rmd
archive earlier, more fragmentary .Rmd's in /legacy/ for clarity

--- 2024-01-16 paper_coding_2_analysis.Rmd
@@
 ---
-title: "Triangulating iconicity: Coding analysis"
+title: "Triangulating iconicity: From structure mapping to guessability"
 author: "Mark Dingemanse & Stella Punselie"
@@
 
-# get consensus coding data
+# get coding data
 d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)
+
+coding_categories <- d %>% dplyr::select(matches('_'),-matches('notes|meaning')) %>% names()
 
@@
 
-# add Z score to make scores more comparable in plots across studies
+# add Z score to account for differences in base success rate of Collabra and Language Studies
 d <- d %>%
@@
 
-# get ratings data
-d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
-  dplyr::select(-category,-list,-item)
-d <- left_join(d,d.ratings,by=c("filename","study","language"))
+```
 
-# add z score for ratings
-d$rating_z <- scale(d$rating,center=T,scale=T)
+We add congruency measures that link F and M elements.
 
-# write this dataset as the fullest dataset
-write.csv(d,file="data/ideophones_coded_guessed_rated.csv")
-write_xlsx(d,path="data/ideophones_coded_guessed_rated.xlsx")
+* TO DISCUSS: We need to think about the evidential value of each of the congruency measures.
+Most apply to many words, which is good.
+If some of them only capture a few (as `C_durative` and `C_weight_tone` seem to do), they are not so interesting because they don't really capture recurring iconic mappings in the data.
+* Discussed: This made us realize a need to name `M_durative` to `M_long`, covering extent not just in time but also in space (a very concrete kind of mapping for `F_finallength`.
+
+```{r congruency_1}
+
+# using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
+d <- d %>%
+  mutate(C_modality = ifelse(M_sound == 1,1,0),
+         C_iterative = ifelse(F_redup == 1 & M_distribution == 1,1,0),
+         C_irregular = ifelse(F_redupmod == 1 & M_irregular == 1,1,0),
+         C_closure = ifelse(F_closedsyllable == 1 & M_abrupt == 1,1,0),
+         C_punctual = ifelse(F_monosyllabic == 1 & M_punctual == 1,1,0),
+         C_long = ifelse(F_finallength == 1 & M_long == 1,1,0),
+         C_weight_voice = ifelse((F_voice == 0 & M_weight == 0) | (F_voice == 2 & M_weight == 2),1,0),
+         C_weight_vowel = ifelse((F_vowelquality == 0 & M_weight == 0) | (F_vowelquality == 2 & M_weight == 2),1,0),
+         C_weight_tone = ifelse((F_intonation == 0 & M_weight == 2) | (F_intonation == 2 & M_weight == 0),1,0)
+         )
+
+# some mappings seem to be mutually exclusive, so it might be fitting to combine them
+d %>%
+  mutate(C_aspect = C_iterative + C_punctual) %>%
+  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_event) %>%
+  filter(C_aspect > 1)
+
+d %>%
+  mutate(C_length = C_long + C_closure) %>%
+  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) %>%
+  filter(C_length > 1)
+
+d <- d %>%
+  mutate(C_aspect = C_iterative + C_punctual,
+         C_length = C_long + C_closure)
+
+# C_weight is recombined as all three form subfeatures are linked to one and the same meaning feature M_weight
+
+d %>% # sanity check
+  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone) %>%
+  dplyr::select(ideophone,meaning_NL,language,C_weight_voice,C_weight_vowel,C_weight_tone,C_weight) %>%
+  filter(C_weight > 0)
+
+d <- d %>%
+  mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
+         C_magnitude = ifelse(C_weight == 0,0,1)) # i.e. a magnitude value that is 1 if any weight feature was congruent
+
+# add cumulative iconicity scores
+d <- d %>%
+  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) %>%
+  mutate(C_any1 = ifelse(C_cumulative == 0,0,1),
+         C_atleast2 = ifelse(C_cumulative > 1,1,0),
+         C_ternary = ifelse(C_cumulative == 0,0,
+                            ifelse(C_cumulative == 1,1,
+                                   ifelse(C_cumulative == 2,2,3)))) # i.e. 1, 2 or 3
+
+# add a simplified cumulative score by summing only the 3 correspondences that account for more than 10% of attested congruent mappings
+
+d <- d %>%
+  mutate(C_simple = C_modality + C_aspect + C_magnitude)
+
+# write this dataset for use in the next step
+write.csv(d,file="data/ideophones_coded_guessed.csv")
 
@@
 d %>%
-  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,rating,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
-  write_xlsx(path="data/ideophones_coded_guessed_rated_simplified.xlsx")
+  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
+  write_xlsx(path="data/ideophones_coded_guessed_simplified.xlsx")
+
+d %>%
+  group_by(C_cumulative) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_any1) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_atleast2) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_ternary) %>%
+  summarise(n=n(),score=mean.na(score))
+
+d %>%
+  group_by(C_simple) %>%
+  summarise(n=n(),score=mean.na(score))
 
@@
 
-## Figures
-Here we generate some descriptive stats and create the main figures for the paper.
+### Figures
+To visualise the results we create the figures for the paper here.
 
-```{r descriptive_stats}
-
-d %>%
-  group_by(study) %>%
-  summarize(max=max(score),min=min(score),mean=mean.na(score))
-
-```
-
-### Figure: Overview panel
-The overview panel is created separately but we use the data from the guessability studies as a canvas.
-
-```{r figures_paper_guessability_canvas}
-
-# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
-
-# blank plot can be useful as a canvas 
-ggplot(data=dp, aes(x=study,y=score)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
-  stat_summary(fun=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
-  geom_dotplot(colour="white",fill="black",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
-
-ggsave("figures\\paper-guessability-blank.pdf",height=5,width=7.5)
-
-d %>% filter(filename=="Korean_23_Motion_org") %>% View()
-
-```
-
-### Figure: Guessability by study and iconicity rating
-
-```{r figures_paper_guessability_rating}
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
-
-# panel A: scores by study and cumulative iconicity
-pA <- ggplot(data=dp, aes(x=study,y=score,fill=study,colour=study)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
-  labs(title="Guessability by study",
-       x="study",
-       y="guessability (% correct)") +
-  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
-  NULL
-
-pB <- ggplot(data=dp, aes(x=rating,y=score,fill=study,colour=study)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") +
-  labs(title="Guessability (both studies) by iconicity rating",
-       x="iconicity rating",
-       y="") +
-  geom_point(size=2.0,shape=21) +
-  geom_smooth(method="lm") +
-  NULL
-
-plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)
-
-```
-
-### Figure: Guessability and cumulative iconicity
-
-```{r figures_paper_guessability_coding}
+```{r figures_paper}
 
@@
   stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
-  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
-  NULL
+  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
 
@@
 
-### Figure: Ratings and cumulative iconicity
-
-```{r ratings_and_C}
-
-# panel A: ratings by study and cumulative iconicity
-pA <- ggplot(data=dp, aes(x="",y=rating,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
-  labs(title="Iconicity ratings for 239 ideophones",
-       x="all ideophones",
-       y="mean iconicity rating") +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-pB <- ggplot(data=dp, aes(x=C_cumulative,y=rating,colour=C_cumulative)) +
-  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
-  labs(title="... by independently coded structural correspondences",
-       x="cumulative iconicity",
-       y="mean iconicity rating") +
-  geom_half_boxplot() +
-  geom_half_point() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  NULL
-
-plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\paper_panelAB_ratings.png",height=5,width=9)
-
-```
-
-### Figure: Mappings and semantic domains
-
-```{r figure_mappings_domains}
-
-library(forcats)
-dp.categories <- d %>%
-  filter(category %notin% "Other") %>%
-  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
-  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
-  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
-  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis
-
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-dp.categories %>%
-  ungroup() %>%
-  group_by(category_binary) %>%
-  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
-group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
-t.test(group1,group2)
-library(effsize)
-cohen.d(group1,group2)
-
-# also when we exclude C_modality? yes
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-dp.categories %>%
-  group_by(category_binary) %>%
-  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
-group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
-t.test(group1,group2)
-cohen.d(group1,group2)
-
-ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
-  theme_tufte() + theme(legend.position="none") +
-  labs(title="Cumulative iconicity by semantic domain") +
-  geom_half_violin() +
-  geom_half_point() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  NULL
-
-ggplot(data=dp.categories, aes(y=C_cumulative,x=category,colour=C_cumulative)) +
-  theme_tufte() + theme(legend.position="none") +
-  labs(title="Cumulative iconicity by semantic domain") +
-  geom_hex() +
-  NULL
-
-```
-
 ### Raincloud plots
@@
 ```{r visuals}
+
+# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
+tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
+dp <- d
+dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
 
@@
   NULL
+ggsave("figures\\coding_guessability_by_cumulative.png",height=5,width=7.5)
 
@@
 
-### Dotplots
+## Dotplots
 
@@
 
-# how do the constituents of C_magnitude fare?
-icoplot("C_magnitude")
-icoplot("C_weight_voice")
-icoplot("C_weight_vowel")
-icoplot("C_weight_tone")
-
 icoplot("C_any1")
@@
 
-# facet by language to discover crosslinguistic diversity
-ggplot(data=dp, aes(x=study,y=score)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
-  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
-  geom_dotplot(colour="white",fill="#c9c9c9",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
+d %>%
+  plot_ly(x = ~logodds, 
+          y = ~C_cumulative,
+          type="scatter3d",
+          mode="markers",
+          color= ~rating_z,
+          name=~ideophone,
+          showlegend=F) %>%
+  layout(showlegend=F)
+
+ggsave("figures\\paper-guessability-blank.eps",height=5,width=7.5)
 
@@
 
-## Numbers
-
-### Coding and guessability
+## Statistical tests
 As the visualizations already make clear there is a non-trivial positive relation between the presence of iconic mappings (in a cumulative sense) and the guessability score determined in experimental work.
@@
 
-### Semantic domains
-Our prediction is that Sound and Motion ideophones should show more structural correspondences than the other domains. This is borne out.
-
-```{r coding_semanticdomains}
-
-library(forcats)
-dp.categories <- d %>%
-  filter(category %notin% "Other") %>%
-  mutate(category = fct_reorder(category, -C_cumulative, .fun='mean')) %>%
-  mutate(category_binary = ifelse(category %in% c("Sound","Motion"),"SoundMotion","ColorvisualShapeTexture")) %>%
-  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude) %>% # for control analysis
-  mutate(C_strongestpredictors = C_aspect + C_length + C_modality ) # for another control analysis
-
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_strongestpredictors),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-dp.categories %>%
-  ungroup() %>%
-  group_by(category_binary) %>%
-  summarise(n=n(),mean_C=mean.na(C_cumulative),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_cumulative
-group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_cumulative
-t.test(group1,group2)
-library(effsize)
-cohen.d(group1,group2)
-
-# also when we exclude C_modality? yes
-dp.categories %>%
-  group_by(category) %>%
-  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-dp.categories %>%
-  group_by(category_binary) %>%
-  summarise(n=n(),mean_C=mean.na(C_nosound),mean_rating=mean.na(rating),mean_score=mean.na(score)) %>%
-  arrange(-mean_C)
-
-group1 <- subset(dp.categories, category_binary == "SoundMotion")$C_nosound
-group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$C_nosound
-t.test(group1,group2)
-cohen.d(group1,group2)
-
-```
-
-We can also look in more detail at the relative differences by semantic domain by building a mixed model that predicts cumulative iconicity with domain as a predictor.
-
-```{r coding_category_lme}
-
-m0 <- lm(C_cumulative ~ category, data=d)
-effsizes(m0)
-
-```
-
-Another prediction is that Sound and Motion ideophones should receive higher iconicity ratings than other domains. This prediction is also borne out.
-
-```{r rating_domains}
-
-group1 <- subset(dp.categories, category_binary == "SoundMotion")$rating
-group2 <- subset(dp.categories, category_binary == "ColorvisualShapeTexture")$rating
-t.test(group1,group2)
-cohen.d(group1,group2)
-
-```
-
-### Coding and ratings
-
-As above, the simplest thing to do is run a simple correlation: there is an r=0.43 correlation between meaning iconicity ratings and cumulative iconicity.
-
-```{r coding_ratings}
-
-cor.test(d$rating_z,d$C_cumulative,method="pearson")
-
-```
-
-As above, we can go a bit deeper by using mixed effects modelling.
-We expect a smaller effect of study on the ratings, as the difference in design in terms of selection of foils falls away.
-However, the studies are also different in terms of
-
-```{r coding_ratings_lme}
-
-m0 <- lmer(rating ~ study + (1|ideophone), data=d)
-m1 <- lmer(rating ~ study + C_cumulative + (1|ideophone), data=d)
-anova(m0,m1)
-
-fixef(m1)
-
-```
-
-We can assess the relative weight of different features using model comparison.
-We build models by adding the congruency measures in rank order of attestation in the coded data.
-Doing this we find that C_aspect, C_modality and C_length are responsible for the bulk of the variance.
-Compared to the null model that only represents the baseline difference between studies, model `m8` fares significantly better (log likelihood difference = 21.3).
-
-The model estimates given in the output of `fixef(m8)` represent the contribution of a given fixed effect while controlling for others, but they are NOT additive (i.e.
-can't be summed to arrive at the effect of `C_cumulative`).
-
-* `C_aspect`: 0.181 logodds (+7% boost)
-* `C_modality`: 0.177 logodds (+7% boost)
-* `C_length`: 0.443 logodds (+17% boost)
-
-```{r congruency_stats_3}
-
-m3 <- lmer(rating ~ study + C_aspect + (1|ideophone), data=d)
-m4 <- lmer(rating ~ study + C_aspect + C_magnitude + (1|ideophone), data=d)
-m5 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + (1|ideophone), data=d)
-m6 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + C_length + (1|ideophone), data=d)
-m7 <- lmer(rating ~ study + C_aspect + C_magnitude + C_modality + C_length + C_irregular + (1|ideophone), data=d)
-m8 <- lmer(rating ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)
-
-anova(m0,m3) # aspect helps
-anova(m3,m4) # magnitude doesn't additionally help
-anova(m4,m5) # modality helps
-anova(m3,m5) # modality helps also relative to m3
-anova(m5,m6) # length helps
-anova(m6,m7) # irregular doesn't additionally help
-anova(m5,m8) # aspect + modality + length is where it's at
-# so m8 is the best performing most complex model
-anova(m0,m8)
-summary(m8)
-fixef(m8)
-
-```
-
-### Control analysis: results stand even when leaving out C_modality
-
-The `C_modality` congruence measure is intended to capture the fact that spoken words can show a range of acoustic similarities to sonic events, independently from other structural correspondences (in other words, it separates imagic iconicity from Gestalt iconicity and relative iconicity).
-We don't code for specific acoustic correspondences, because these would be by nature only apply to ideophones evocative of sound and would greatly complicate the coding scheme.
-
-This makes it the least fine-grained coding category.
-And given that it applies to a semantic domain that is known, from prior work and first principles, to be easiest to guess, one could think of `C_modality` as just giving sound words a leg up in the cumulative scores.
-Therefore, as a control, it is useful to know whether the patterns we find hold up even if we disregard `C_modality` entirely.
-
-```{r sound_off}
-
-# we add C_nosound as a cumulative measure that excludes modality
-d <- d %>%
-  mutate(C_nosound = C_aspect + C_length + C_irregular + C_magnitude)
-
-# recreate dp to include C_nosound
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
-
-d %>%
-  group_by(C_nosound) %>%
-  summarise(n=n(),score=mean.na(score))
-
-ggplot(data=dp, aes(x=C_nosound,y=score_z,colour=C_nosound)) +
-  theme_tufte() + theme(legend.position="none") +
-  labs(title="Guessability (both studies) by cumulative iconicity",
-       x="cumulative iconicity (excluding modality)",
-       y=expression('guessability ('~italic(z)~')')) +
-  geom_half_boxplot() +
-  geom_half_point() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  NULL
-
-```
-
-There is an r=0.26 correlation between logodds and C_sound across the studies. This is present in both studies.
-
-```{r congruency_stats_nosound_1}
-
-# simplest thing we can do is a correlation.
-
-cor.test(d$logodds,d$C_nosound,method="pearson")
-
-d.Collabra <- d %>% filter(study == "Collabra")
-d.Language <- d %>% filter(study == "Language")
-cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")
-cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")
-
-p0 <- cor.test(d$logodds,d$C_nosound,method="pearson")$p.value
-p1 <- cor.test(d.Collabra$logodds,d.Collabra$C_nosound,method="pearson")$p.value
-p2 <- cor.test(d.Language$logodds,d.Language$C_nosound,method="pearson")$p.value
-p.adjust(c(p0,p1,p2),method="bonferroni")
-
-```
-
-The mixed effects models look much the same.
-A null model `m0` predices logodds by paper, controlling for item-level differences. A model `m1` with cumulative congruence as an additional fixed effect fares significantly better.
-
-For the fixed effects, m1 lists a 0.205 improvement in logodds for `C_nosound`.
-According to my interpretation this means that every added iconic feature leads to an ~8% boost in proportion correct.
-The lower baseline for the Language set is estimated as -0.513, which translates to a 19.5% guessability penalty.
-
-```{r congruency_stats_nosound2}
-
-m0 <- lmer(logodds ~ study + (1|ideophone), data=d)
-m1 <- lmer(logodds ~ study + C_nosound + (1|ideophone), data=d)
-anova(m0,m1)
-
-fixef(m1)
-
-probitlink(0.205,inverse=T)-0.5
-probitlink(-0.513,inverse=T)-0.5
-
-```
-
-## Exploring coding sensitivity and congruency blind spots
+## Exploring congruency blind spots
 What are we not capturing?
@@
 
-Coding sensitivity varies by feature, in part because not all are equally frequent.
-For instance, `C_weight_tone` captures ideophones that have high tone/intonation and small/light meaning or conversely low tone/intonation and big/heavy meaning.
-
-```{r coding_sensitivity}
-
-icoplot("C_weight_tone")
-
-d %>%
-  group_by(language) %>%
-  filter(C_weight_tone == 1) %>%
-  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))
-
-d %>%
-  group_by(language) %>%
-  filter(C_weight_tone == 1) %>%
-  View()
-
-```
-
-### Variation by language
-
-Although we have focused on cross-linguistically attested features, features may also vary by language.
-For instance, irregularity may work somewhat in Japanese, Korean and Ewe, the 4 Semai ideophones scored as showing this mapping have a mean accuracy score of 0.45, dragging down the average.
-It looks like partial reduplication really has a different function in Semai, and one that is not captured by our coding scheme.
-
-```{r coding_by_language}
-
-d %>%
-  group_by(language) %>%
-  filter(C_irregular == 1) %>%
-  summarise(n=n(),score=mean.na(score),rating=mean.na(rating),C=mean.na(C_cumulative))
-
-d %>%
-  group_by(language) %>%
-  filter(C_irregular == 1) %>%
-  dplyr::select(ideophone,meaning,language,category,score,rating,C_irregular,C_cumulative) %>%
-  View()
-
-mS0 <- lmer(logodds ~ study + (1|ideophone), data=d)
-mS1 <- lmer(logodds ~ study + language + (1|ideophone), data=d) # fails to converge
-mS2 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + (1|ideophone), data=d)
-mS3 <- lmer(logodds ~ study + C_aspect + C_modality + C_length + language + (1|ideophone), data=d)
-
-anova(mS2,mS3)
-
-fixef(mS3)
-
-```
-
-One might conclude from that language would be a relevant factor to take into account in the statistical models.
-But its influence is not very large.
-If we add it to the most complete model above, the fit is marginally improved, but the difference is not of the same magnitude as other factors.
-Also, language is partly confounded with quality of recordings, which are lab recordings for Ewe, Korean, Japanese and field recordings for Siwu and Semai.
-We can dummy code this as a new variable to check its effect.
-
-```{r coding_by_audio}
-
-d <- d %>%
-  mutate(audio = ifelse(language %in% c("Siwu","Semai"),"field","lab"))
-
-d %>%
-  group_by(audio) %>%
-  summarise(n=n(),score=mean.na(score),rating=mean.na(rating))
-
-ggplot(data=d, aes(x=audio,y=score,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
-  labs(title="Guessability by audio version",
-       x="audio",
-       y="guessability (% correct)") +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-ggplot(data=d, aes(x=language,y=score,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
-  labs(title="Guessability by audio version",
-       x="language",
-       y="guessability (% correct)") +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-# ratings look like they have a distinctly non-normal distribution for field recordings, with a large bulge below 3 (= middle)
-ggplot(data=d, aes(x=audio,y=rating,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
-  labs(title="Iconicity ratings for 239 ideophones",
-       y="mean iconicity rating") +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-# and indeed the bulge isn't really there for any of the non-field languages (and arguably mostly attributable to the Semai recordings)
-ggplot(data=d, aes(x=language,y=rating,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
-  labs(title="Iconicity ratings for 239 ideophones",
-       x="language",
-       y="mean iconicity rating") +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-```
-
 ```{r congruency_underperformers}

### 2024-04-03 legacy/TI_03_ratings.Rmd
archive earlier, more fragmentary .Rmd's in /legacy/ for clarity

--- 2024-01-16 TI_03_ratings.Rmd
@@
 # load corrected data
-d <- read_xlsx("data\\ideophones_rated.xlsx")
-d.means <- read_xlsx("data\\ideophones_rated_means.xlsx")
+d.allratings <- read_xlsx("data\\ideophones_rated.xlsx")
+d.allratings.means <- read_xlsx("data\\ideophones_rated_means.xlsx")
 
 # get means and standard deviations across all ideophones and for each language and category
-mean(d$rating)
+mean(d.allratings$rating)
   # 2.948339
-sd(d$rating)
+sd(d.allratings$rating)
   # 1.297106
-sumcat <- ddply(d,~category,summarise,mean=mean(rating),sd=sd(rating))
-sumlang <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))
+sumcat <- ddply(d.allratings,~category,summarise,mean=mean(rating),sd=sd(rating))
+sumlang <- ddply(d.allratings,~language,summarise,mean=mean(rating),sd=sd(rating))
 
@@
 #create copy of data including the category "Other"
-d.other <- read_xlsx("data\\ideophones_rated.xlsx")
-d.m.other <- read_xlsx("data\\ideophones_rated_means.xlsx")
+d.allratings.other <- read_xlsx("data\\ideophones_rated.xlsx")
+d.allratings.means.other <- read_xlsx("data\\ideophones_rated_means.xlsx")
 
 #remove category "Other" from d and d.means
-d <- d[!grepl("Other",d$category),]
-d.means <- d.means[!grepl("Other",d.means$category),]
+d.allratings <- d.allratings[!grepl("Other",d.allratings$category),]
+d.allratings.means <- d.allratings.means[!grepl("Other",d.allratings.means$category),]
 
 # make new summary for language (without the Japanese "other" ideophones)
-sumlang2 <- ddply(d,~language,summarise,mean=mean(rating),sd=sd(rating))
+sumlang2 <- ddply(d.allratings,~language,summarise,mean=mean(rating),sd=sd(rating))
 ```
@@
 cols <- c("category","language","study")
-d[cols] <- lapply(d[cols], factor)
-d.means[cols] <- lapply(d.means[cols], factor)
+d.allratings[cols] <- lapply(d.allratings[cols], factor)
+d.allratings.means[cols] <- lapply(d.allratings.means[cols], factor)
 
 #combined boxplot and scatterplot by category
-ggplot(d.means, aes(x=reorder(category,-rating),y=rating,color=category)) +
+ggplot(d.allratings.means, aes(x=reorder(category,-rating),y=rating,color=category)) +
   theme_tufte(base_size=16) +
@@
 #same, but by language
-ggplot(d.means, aes(x=reorder(language,-rating),y=rating,color=language)) +
+ggplot(d.allratings.means, aes(x=reorder(language,-rating),y=rating,color=language)) +
   theme_tufte(base_size=16) +
@@
 # check normality by visual inspection
-ggqqplot(d, "rating", facet.by = "category")
+ggpubr::ggqqplot(d.allratings, "rating", facet.by = "category")
   # all points fall approximately along the reference line, so normality can be assumed
@@
 ```{r}
-summary(lm.cat <- lm(rating ~ category, d))
+summary(lm.cat <- lm(rating ~ category, d.allratings))
 
 # pairwise comparisons with Tukey adjustments in the emmeans package
-emm.cat <- emmeans(lm.cat, "category")
+emm.cat <- emmeans::emmeans(lm.cat, "category")
 pairs(emm.cat)
@@
 ```{r}
-summary(lm.lang <- lm(rating ~ language, d))
+summary(lm.lang <- lm(rating ~ language, d.allratings))
 

### 2024-04-03 legacy/TI_04_triangulation.Rmd
archive earlier, more fragmentary .Rmd's in /legacy/ for clarity

---
  title: "Triangulating iconicity: combining coding, guessing and rating measures"
author: "Mark Dingemanse & Stella Punselie"
date: "Updated `r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

  ```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='figures_md/',
                      echo=TRUE, warning=FALSE, message=FALSE)

```

Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores.

## Setup

```{r preliminaries, results="hide"}
# Packages
list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","gghalves","plotly","viridis")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
lapply(list.of.packages, require, character.only=T)

# useful functions
`%notin%` <- function(x,y) !(x %in% y) 
mean.na <- function(x) mean(x, na.rm = T)
sd.na <- function(x) sd(x, na.rm = T)

```
## Data

```{r data}
d = read.csv("data\\ideophones_coded_guessed_rated.csv")

coding_categories <- d %>% dplyr::select(matches('F_|M_'),-matches('notes|meaning')) %>% names()

# add cumulative score to allow summing of cumulative, rating, guessability
d$cumulative_z <- scale(d$C_cumulative,center=T,scale=T)
d <- d %>%
  mutate(overallrank = as.numeric(cumulative_z + rating_z + score_z))

d %>% dplyr::select(ideophone,
                    meaning,
                    meaning_NL,
                    category,
                    language,
                    study,
                    score,
                    rating,
                    C_cumulative,
                    overallrank) %>%
  filter(score > .65,
         category == "Motion",
         C_cumulative > 0) %>%
  dplyr::arrange(-overallrank) %>%
  View()

# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
dp <- d
dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)

```

## Visualizations
Let's have a look at the relations between the coding, guessing and rating data.
```{r visuals}

ggplot(d,aes(x=rating_z,y=logodds, color=C_cumulative)) +
  geom_point() +
  geom_smooth(method=loess,color="black") +
  scale_fill_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
  scale_colour_viridis(option="plasma",discrete=F,begin=0.3,end=0.9) +
  NULL

d %>%
  plot_ly(x = ~rating_z, 
          y = ~logodds, 
          z = ~C_cumulative,
          type="scatter3d",
          mode="markers",
          color= ~rating_z,
          name=~ideophone,
          showlegend=F) %>%
  layout(showlegend=F)

d %>%
  plot_ly(x = ~rating_z, 
          y = ~logodds, 
          type= "scatter",
          color= ~rating_z,
          name=~ideophone,
          showlegend=F) %>%
  layout(showlegend=F)

ggplot(dp, aes(x=C_cumulative,y=rating_z,fill=score_z,color=score_z)) +
  theme_tufte() +
  geom_half_boxplot() +
  geom_half_point() +
  NULL

## look at some data

d %>%
  filter(rating_z >1,
         logodds >0.6,
         C_cumulative >1) %>%
  View()

```

### 2024-04-03 legacy/TI_05_plotting_explorations.Rmd
archive earlier, more fragmentary .Rmd's in /legacy/ for clarity

--- 2024-01-16 paper_coding_2_analysis.Rmd
@@
 # get ratings data
-d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
+d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") %>%
   dplyr::select(-category,-list,-item)
@@
 d$rating_z <- scale(d$rating,center=T,scale=T)
-
-# write this dataset as the fullest dataset
-write.csv(d,file="data/ideophones_coded_guessed_rated.csv")
-write_xlsx(d,path="data/ideophones_coded_guessed_rated.xlsx")
-
-# also write a simplified copy for easy inspection in Excel (helpful when writing)
-d %>%
-  dplyr::select(ideophone,meaning,language,category,meaning_NL,study,score,rating,C_aspect,C_magnitude,C_modality,C_length,C_irregular,C_cumulative) %>%
-  write_xlsx(path="data/ideophones_coded_guessed_rated_simplified.xlsx")
-
-```
-
-## Figures
-Here we generate some descriptive stats and create the main figures for the paper.
-
-```{r descriptive_stats}
-
-d %>%
-  group_by(study) %>%
-  summarize(max=max(score),min=min(score),mean=mean.na(score))
-
-```
-
-### Figure: Overview panel
-The overview panel is created separately but we use the data from the guessability studies as a canvas.
-
-```{r figures_paper_guessability_canvas}
-
-# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
-
-# blank plot can be useful as a canvas 
-ggplot(data=dp, aes(x=study,y=score)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none",axis.title.x=element_blank(),plot.margin=margin(0,0,10,0)) + 
-  stat_summary(fun=median,geom="point",size=8,shape=21,stroke=1,fill="white",colour="#c9c9c9") +
-  geom_dotplot(colour="white",fill="black",stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center")
-
-ggsave("figures\\paper-guessability-blank.pdf",height=5,width=7.5)
-
-d %>% filter(filename=="Korean_23_Motion_org") %>% View()
-
-```
-
-### Figure: Guessability by study and iconicity rating
-
-```{r figures_paper_guessability_rating}
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
-
-# panel A: scores by study and cumulative iconicity
-pA <- ggplot(data=dp, aes(x=study,y=score,fill=study,colour=study)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
-  labs(title="Guessability by study",
-       x="study",
-       y="guessability (% correct)") +
-  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
-  NULL
-
-pB <- ggplot(data=dp, aes(x=rating,y=score,fill=study,colour=study)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") +
-  labs(title="Guessability (both studies) by iconicity rating",
-       x="iconicity rating",
-       y="") +
-  geom_point(size=2.0,shape=21) +
-  geom_smooth(method="lm") +
-  NULL
-
-plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\paper_panelAB_guessing_rating.png",height=5,width=9)
-
-```
-
-### Figure: Guessability and cumulative iconicity
-
-```{r figures_paper_guessability_coding}
-
-# make a copy dp of the data in which we convert some measures to factors for easy plotting & facetting
-tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
-dp <- d
-dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
-
-# panel A: scores by study and cumulative iconicity
-pA <- ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
-  labs(title="Guessability by study",
-       x="study",
-       y="guessability (% correct)") +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
-  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
-  NULL
-
-pB <- ggplot(data=dp, aes(x=C_cumulative,y=score_z,colour=C_cumulative)) +
-  theme_tufte() + theme(legend.position="none") +
-  labs(title="Guessability (both studies) by cumulative iconicity",
-       x="cumulative iconicity",
-       y=expression('guessability ('~italic(z)~')')) +
-  geom_half_boxplot() +
-  geom_half_point() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  NULL
-
-plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\paper_panelAB.png",height=5,width=9)
-
-```
-
-### Figure: Ratings and cumulative iconicity
-
-```{r ratings_and_C}
-
-# panel A: ratings by study and cumulative iconicity
-pA <- ggplot(data=dp, aes(x="",y=rating,fill=C_cumulative,colour=C_cumulative)) +
-  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
-  labs(title="Iconicity ratings for 239 ideophones",
-       x="all ideophones",
-       y="mean iconicity rating") +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  geom_beeswarm(cex=2.5) +
-  NULL
-
-pB <- ggplot(data=dp, aes(x=C_cumulative,y=rating,colour=C_cumulative)) +
-  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
-  labs(title="... by independently coded structural correspondences",
-       x="cumulative iconicity",
-       y="mean iconicity rating") +
-  geom_half_boxplot() +
-  geom_half_point() +
-  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
-  NULL
-
-plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures\\paper_panelAB_ratings.png",height=5,width=9)
 

### 2024-04-03 paper_03_exploratory_analyses.R
use native pipe; move new .tsv to /data

@@ -7,41 +7,41 @@ library(viridis)
 
 #### So I just make it again here because that works
 # get consensus coding data
-d = read_excel("data\\ideophones_coded.xlsx") %>% arrange(filename)
+d = read_excel("data/ideophones_coded.xlsx") |> arrange(filename)
 
 # add guessability scores from the Collabra and Language studies.
 
-d.scores = read_excel("data\\ideophones_guessability.xlsx") %>%
+d.scores = read_excel("data/ideophones_guessability.xlsx") |>
   dplyr::select(-category)
 d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))
 
 
 # add Z score to make scores more comparable in plots across studies
-d <- d %>%
-  group_by(study) %>%
+d <- d |>
+  group_by(study) |>
   mutate(score_z = scale(score,center=T,scale=T))
 
 # get ratings data
-d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") %>%
+d.ratings <- read_xlsx("data\\ideophones_rated_means.xlsx") |>
   dplyr::select(-category,-list,-item)
 d <- left_join(d,d.ratings,by=c("filename","study","language"))
 
 # add z score for ratings
 d$rating_z <- scale(d$rating,center=T,scale=T)
 
-d%>%
-  select(-C_length,-C_aspect,-C_magnitude,-C_weight)%>%
-  pivot_longer(C_modality:C_weight_tone,names_to="correlate",values_to="present")%>%
-  filter(present!=0)%>%
-  mutate(features="features")%>%
-  group_by(ideophone)%>%
-  pivot_wider(names_from="features",values_from="correlate",values_fn=list)%>%
+d|>
+  select(-C_length,-C_aspect,-C_magnitude,-C_weight) |>
+  pivot_longer(C_modality:C_weight_tone,names_to="correlate",values_to="present") |>
+  filter(present!=0) |>
+  mutate(features="features") |>
+  group_by(ideophone) |>
+  pivot_wider(names_from="features",values_from="correlate",values_fn=list) |>
   mutate(label=paste(ideophone,meaning,features,sep="\n"))->dat
 
 # Overview plot
 
 dat$category <- factor(dat$category,levels=c("Sound","Motion","Shape","Texture","ColorVisual","Other"))
-dat%>%
+dat |>
[…]

### 2024-04-03 app.R
move

@@ -80,7 +80,7 @@ ui <- fluidPage(
       
     ),
   ),
-  fluidRow(includeHTML("codingscheme.html"))
+  fluidRow(includeHTML("data/codingscheme.html"))
 )
 
 server <- function(input, output) {

### 2024-04-03 paper_01_consistency.Rmd
use native pipe & check whether everything still renders

@@
 ---
-title: "Triangulating iconicity: Coding, consistency & calibration"
-author: "Mark Dingemanse & Stella Punselie"
+title: "Coding, consistency & calibration"
+author: "[anonymized for review]"
 date: "Updated `r format(Sys.Date())`"
@@
 list.of.packages <- c("tidyverse","readxl","writexl","irr")
-new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
-if(length(new.packages)) install.packages(new.packages)
+# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
+# if(length(new.packages)) install.packages(new.packages)
 lapply(list.of.packages, require, character.only=T)
@@
 
-d1 = read_excel("data/ideophones_coded_MD_original.xlsx") %>% arrange(filename)
-d2 = read_excel("data/ideophones_coded_SP_original.xlsx") %>% arrange(filename)
+d1 = read_excel("data/ideophones_coded_MD_original.xlsx") |> arrange(filename)
+d2 = read_excel("data/ideophones_coded_SP_original.xlsx") |> arrange(filename)
 d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge
@@
 
-d.F <- d %>%
-  dplyr::select(c("filename","ideophone"),starts_with("F_")) %>%
+d.F <- d |>
+  dplyr::select(c("filename","ideophone"),starts_with("F_")) |>
   dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
@@
 
-d.M <- d %>%
-  dplyr::select(c("filename","ideophone"),starts_with("M_")) %>%
+d.M <- d |>
+  dplyr::select(c("filename","ideophone"),starts_with("M_")) |>
   dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
@@
 get_coding_categories <- function(data) {
-  coder_1_names <- data %>% dplyr::select(ends_with("_MD")) %>% names()
+  coder_1_names <- data |> dplyr::select(ends_with("_MD")) |> names()
   gsub('MD','',coder_1_names)
@@
 F_categories <- get_coding_categories(d.F)
-F_consistency <- as.data.frame(F_categories) %>% dplyr::rename(category = F_categories)
+F_consistency <- as.data.frame(F_categories) |> dplyr::rename(category = F_categories)
 M_categories <- get_coding_categories(d.M)
-M_consistency <- as.data.frame(M_categories) %>% dplyr::rename(category = M_categories)
+M_consistency <- as.data.frame(M_categories) |> dplyr::rename(category = M_categories)
 
@@
 ```{r consistency_2ndround}
-d1 = read_excel("data/ideophones_coded_MD.xlsx") %>% arrange(filename)
-d2 = read_excel("data/ideophones_coded_SP.xlsx") %>% arrange(filename)
+d1 = read_excel("data/ideophones_coded_MD.xlsx") |> arrange(filename)
+d2 = read_excel("data/ideophones_coded_SP.xlsx") |> arrange(filename)
 d2 <- d2[,-c(1:7)] # remove non-unique fields except filename prior to merge
@@
 
-d.F <- d %>%
-  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("F_")) %>%
+d.F <- d |>
+  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("F_")) |>
   dplyr::select(-starts_with("F_notes"),everything()) # put note fields at end
 
-d.M <- d %>%
-  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("M_")) %>%
+d.M <- d |>
+  dplyr::select(c("filename","ideophone","meaning","meaning_NL"),starts_with("M_")) |>
   dplyr::select(-starts_with("M_notes"),everything()) # put note fields at end
@@
 inconsistent <- c('F_redup','F_monosyllabic','F_vowelquality','F_finallength')
-d.F.inconsistent <- d.F %>%
-  dplyr::select(filename,ideophone,meaning,meaning_NL,matches(paste(inconsistent,collapse="|"))) %>%
+d.F.inconsistent <- d.F |>
+  dplyr::select(filename,ideophone,meaning,meaning_NL,matches(paste(inconsistent,collapse="|"))) |>
   mutate(F_redup_C = ifelse(F_redup_MD == F_redup_SP,1,0),
@@
          F_vowelquality_C = ifelse(F_vowelquality_MD == F_vowelquality_SP,1,0),
-         F_finallength_C = ifelse(F_finallength_MD == F_finallength_SP,1,0)) %>%
-  mutate(check = F_redup_C + F_monosyllabic_C + F_vowelquality_C + F_finallength_C) %>%
+         F_finallength_C = ifelse(F_finallength_MD == F_finallength_SP,1,0)) |>
+  mutate(check = F_redup_C + F_monosyllabic_C + F_vowelquality_C + F_finallength_C) |>
   filter(check < 4)
@@
 
-d.coded <- read_excel("data/ideophones_coded_SP.xlsx") %>% 
-  arrange(filename) %>%
+d.coded <- read_excel("data/ideophones_coded_SP.xlsx") |> 
+  arrange(filename) |>
   dplyr::select(-matches('notes'))
@@
 
-notes <- d %>% dplyr::select(filename,matches('notes'))
+notes <- d |> dplyr::select(filename,matches('notes'))
 
@@
 # using ifelse() statements that essentially say, if these conditions are true, use 1, otherwise 0
-d <- d.coded %>%
+d <- d.coded |>
   mutate(C_modality = ifelse(M_sound == 1,1,0),
@@
 #sanity check: are they indeed mutually exclusive? these calls should come up empty
-d %>% 
-  mutate(C_aspect = C_iterative + C_punctual) %>%
-  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_aspect) %>%
+d |> 
+  mutate(C_aspect = C_iterative + C_punctual) |>
+  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_punctual,C_aspect) |>
   filter(C_aspect > 1)
-d %>%
-  mutate(C_length = C_long + C_closure) %>%
-  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) %>%
+d |>
+  mutate(C_length = C_long + C_closure) |>
+  dplyr::select(ideophone,meaning_NL,language,C_iterative,C_closure,C_length) |>
   filter(C_length > 1)
 
-d <- d %>%
+d <- d |>
   mutate(C_aspect = C_iterative + C_punctual,
@@
 
-d <- d %>%
+d <- d |>
   mutate(C_weight = C_weight_voice + C_weight_vowel + C_weight_tone,
@@
 # ternary = simple way of conflating the (smaller) upper regions: 1=1, 2=2, >2 = 3
-d <- d %>%
-  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) %>%
+d <- d |>
+  mutate(C_cumulative = C_modality + C_aspect + C_length + C_irregular + C_magnitude) |>
   mutate(C_any1 = ifelse(C_cumulative == 0,0,1),

### 2024-04-27 paper_02_main_analyses.Rmd
+csv version of full ratings data
anon

@@
 title: "Triangulating iconicity: Coding analysis"
-author: "Mark Dingemanse, Stella Punselie, Bonnie McLean"
+author: "[anonymised for review]"
 date: "Updated `r format(Sys.Date())`"
@@
                   ideophones_filename,by="filename",copy=TRUE)
+
+write.csv(d.ratings.full, file="data/ideophones_rated.csv")
 

### 2024-06-15 paper_02_main_analyses.Rmd
separate data processing from analysis

--- 2024-04-03 paper_02_main_analyses.Rmd
@@
 # Packages
-list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","gghalves","ggbeeswarm","viridis","lme4","VGAM","cowplot")
+list.of.packages <- c("tidyverse","ggthemes","gghalves","ggbeeswarm","viridis","lme4","VGAM","cowplot")
 new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
@@
 ## Data
-Here we load the ground truth version of the coded data and add the independently collected guessability scores from the 2016 Collabra and Language papers.
+Here we load the data. For earlier processing operations, see `paper_00_data_processing.R`.
 
-```{r data}
+```{r load_data}
+d <- readr::read_csv("data/ideophones_coded_guessed_rated.csv")
 
-# get consensus coding data
-d = read_excel("data/ideophones_coded.xlsx") |> arrange(filename)
-
-# add guessability scores from the Collabra and Language studies.
-
-d.scores = read_excel("data/ideophones_guessability.xlsx") |>
-  dplyr::select(-category)
-d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))
-
-# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
-d <- d |>
-  group_by(study) |>
-  mutate(logodds = probitlink(score))
-
-# add Z score to make scores more comparable in plots across studies
-d <- d |>
-  group_by(study) |>
-  mutate(score_z = scale(score,center=T,scale=T))
-
-# get ratings data
-d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") |>
-  dplyr::select(-category,-list,-item)
-d <- left_join(d,d.ratings,by=c("filename","study","language")) |>
-  mutate(filename = gsub("_org","",filename))
-
-# add z score for ratings
-d$rating_z <- scale(d$rating,center=T,scale=T)
-
-# write this dataset as the fullest dataset
-write.csv(d,file="data/ideophones_coded_guessed_rated.csv",fileEncoding = "UTF-8")
-write_xlsx(d,path="data/ideophones_coded_guessed_rated.xlsx")
-
-# load and merge the full ratings data for some of the reporting in the paper
-ideophones_filename <- d |> ungroup() |> select(ideophone,filename) |>
-  mutate(filename = str_to_lower(filename))
-
-d.ratings.full <- read_xlsx("data/ideophones_rated.xlsx") |>
-  mutate(list = gsub(".*([A-Z])$", "\\1",item)) |>
-  mutate(filename = gsub("^(.*)_.*$", "\\1", str_to_lower(item))) |>
-  mutate(filename = gsub("Coll_","", filename)) |>
-  select(-item)
-
-# match the data to ideophones for easier lookup
-d.ratings.full <- left_join(d.ratings.full,
-                  ideophones_filename,by="filename",copy=TRUE)
+d.ratings.full <- readr::read_csv("data/ideophones_rated.csv")
 

### 2024-06-15 paper_00_data_processing.R

@@ -1,0 +1,70 @@
+# Data processing for Triangulating iconicity
+
+# Packages
+list.of.packages <- c("tidyverse","readxl","writexl","ggthemes","gghalves","ggbeeswarm","viridis","lme4","VGAM","cowplot")
+new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages)
+lapply(list.of.packages, require, character.only=T)
+
+# useful functions
+`%notin%` <- function(x,y) !(x %in% y) 
+mean.na <- function(x) mean(x, na.rm = T)
+sd.na <- function(x) sd(x, na.rm = T)
+
+
+# Load data ---------------------------------------------------------------
+
+
+#Here we load the ground truth version of the coded data and add the independently collected guessability scores from the 2016 Collabra and Language papers.
+
+# get consensus coding data
+d = read_excel("data/ideophones_coded.xlsx") |> arrange(filename)
+
+# add guessability scores from the Collabra and Language studies.
+
+d.scores = read_excel("data/ideophones_guessability.xlsx") |>
+  dplyr::select(-category)
+d <- left_join(d,d.scores,by=c("ideophone","language","study" = "paper"))
+
+# add logodds (for when we run stats: it's more sensible to predict against logodds than raw proportion correct)
+d <- d |>
+  group_by(study) |>
+  mutate(logodds = probitlink(score))
+
+# add Z score to make scores more comparable in plots across studies
+d <- d |>
+  group_by(study) |>
+  mutate(score_z = scale(score,center=T,scale=T))
+
+# get ratings data
+d.ratings <- read_xlsx("data/ideophones_rated_means.xlsx") |>
+  dplyr::select(-category,-list,-item)
+d <- left_join(d,d.ratings,by=c("filename","study","language")) |>
+  mutate(filename = gsub("_org","",filename))
+
+# add z score for ratings
+d$rating_z <- scale(d$rating,center=T,scale=T)
+
+
+# Write data --------------------------------------------------------------
+
+
+# write this dataset as the fullest dataset
+write.csv(d,file="data/ideophones_coded_guessed_rated.csv",fileEncoding = "UTF-8",row.names=FALSE)
+write_xlsx(d,path="data/ideophones_coded_guessed_rated.xlsx")
+
[…]

### 2024-06-15 paper_02_main_analyses.Rmd
knit 'n commit

@@
 
+# variant with shadow for use in paper
+# panel A: scores by study and cumulative iconicity
+pA <- ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
+  theme_tufte() + ylim(0,1) + theme(legend.position="none") + 
+  labs(title="Guessability by study",
+       x="study",
+       y="guessability (% correct)") +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
+  ggfx::with_shadow(
+    geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center"),
+    sigma=0,
+    x_offset=1.4,
+    y_offset=1.4,
+    colour="black") +
+  NULL
+
+pB <- ggplot(data=dp, aes(x=C_cumulative,y=score_z,colour=C_cumulative)) +
+  theme_tufte() + theme(legend.position="none") +
+  labs(title="Guessability (both studies) by cumulative iconicity",
+       x="cumulative iconicity",
+       y=expression('guessability ('~italic(z)~')')) +
+  geom_half_boxplot() +
+  ggfx::with_shadow(
+    geom_half_point(),
+    sigma=0,
+    x_offset=1.4,
+    y_offset=1.4,
+    colour="black") +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  NULL
+
+#plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
+#ggsave("figures/fig3-panelAB-shadow.png",height=4,width=9,bg="white")
+
 ```
@@
   geom_half_point(show.legend=F) +
-  scale_colour_viridis_d(option="D",alpha=0.8) +
+  scale_colour_viridis_d(option="D") +
   xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
@@
 ggsave("figures/fig4-panel_ratings.png",height=4,width=9,bg="white")
+
+# variant with shadow for possible use in paper
+
+#panel A: ratings by study
+pA <- d |>  ggplot(aes(x=score_z,y=rating)) +   
+  theme_tufte(base_size = 16) +
+  theme(legend.position = c(0.2,0.9),
+        legend.title = element_blank()) +
+  scale_fill_manual(values=c("white","black")) +
+  geom_point(aes(fill=study),
+             position="jitter",
+             shape=21,
+             colour="black") +
+  geom_smooth(method=loess,colour="black",alpha=0.5) +
+  xlab("guessability (z)") + ylab("rating") +
+  NULL
+
+#panel B: ratings by semantic domain
+pB <- d |>
+  filter(category != "Other") |>
+  ggplot(aes(x=reorder(category,-rating),y=rating,color=category)) +
+  theme_tufte(base_size=16) +
+  geom_half_boxplot(aes(middle=mean(rating)),show.legend=F) +
+  ggfx::with_shadow(
+    geom_half_point(show.legend=F),
+    sigma=0,
+    x_offset=1.4,
+    y_offset=1.4,
+    colour="black"
+    ) +
+  scale_colour_viridis_d(option="D") +
+  xlab("category") + scale_x_discrete(labels = c("Sound","Motion","Shape", 
+  "Texture","Colour/Visual"))
+
+#plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.6,2))
+#ggsave("figures/fig4-panel_ratings-shadow.png",height=4,width=9,bg="white")
 
@@
 
-### Figure 5: Ratings and cumulative iconicity
+### Figure 6: Ratings and cumulative iconicity
 
@@
 # panel A: ratings by study and cumulative iconicity
-pA <- ggplot(data=dp, aes(x="",y=rating,fill=C_cumulative,colour=C_cumulative)) +
+pA <- ggplot(data=dp, 
+             aes(x="",
+                 y=rating,
+                 colour=C_cumulative,
+                 fill=C_cumulative)) +
   theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
@@
 
-pB <- ggplot(data=dp, aes(x=C_cumulative,y=rating,colour=C_cumulative)) +
+pB <- ggplot(data=dp, 
+             aes(x=C_cumulative,
+                 y=rating,
+                 colour=C_cumulative,
+                 fill=C_cumulative)) +
   theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
@@
 ggsave("figures/fig6-panelAB_ratings.png",height=5,width=9,bg="white")
+
+```
+
+R1 asks for a variant of this figure where the dots have black outline, to make the most yellow dots more easily visible. A second variant with subtle dropshadow may be even better.
+
+```{r R1_request}
+
+pA <- ggplot(data=dp, 
+             aes(x="",
+                 y=rating,
+                 colour=C_cumulative,
+                 fill=C_cumulative)) +
+  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
+  labs(title="Iconicity ratings for 239 ideophones",
+       x="all ideophones",
+       y="mean iconicity rating") +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  geom_beeswarm(cex=2.5,
+                shape=21,
+                colour="black") +
+  NULL
+
+pB <- ggplot(data=dp, 
+             aes(x=C_cumulative,
+                 y=rating,
+#                 colour=C_cumulative,
+                 fill=C_cumulative)) +
+  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
+  labs(title="... by independently coded structural correspondences",
+       x="cumulative iconicity",
+       y="mean iconicity rating") +
+  geom_half_boxplot() +
+  geom_half_point(shape=21,colour="black") +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  NULL
+
+plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
+ggsave("figures/fig6-panelAB_ratings_variant1.png",height=5,width=9,bg="white")
+
+# what about a drop shadow?
+
+pA <- ggplot(data=dp, 
+             aes(x="",
+                 y=rating,
+                 colour=C_cumulative,
+                 fill=C_cumulative)) +
+  theme_tufte() +  ylim(1,5) + theme(legend.position="none") + 
+  labs(title="Iconicity ratings for 239 ideophones",
+       x="all ideophones",
+       y="mean iconicity rating") +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  ggfx::with_shadow(
+    geom_beeswarm(cex=2.5),
+    sigma=0,
+    x_offset=1.4,
+    y_offset=1.4,
+    colour="black"
+  ) +
+  NULL
+
+pB <- ggplot(data=dp, 
+             aes(x=C_cumulative,
+                 y=rating)) +
+  theme_tufte() +  ylim(1,5) + theme(legend.position="none") +
+  labs(title="... by independently coded structural correspondences",
+       x="cumulative iconicity",
+       y="mean iconicity rating") +
+  geom_half_boxplot(aes(colour=C_cumulative)) +
+  ggfx::with_shadow(
+    geom_half_point(aes(colour=C_cumulative,fill=C_cumulative),
+                    shape=21),
+    sigma=0,
+    x_offset=1.4,
+    y_offset=1.4,
+    colour="black") +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  NULL
+
+plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
+ggsave("figures/fig6-panelAB_ratings_variant2.png",height=5,width=9,bg="white")
 

### 2024-09-12 paper_02_main_analyses.Rmd
+publication quality plots

@@
 
-### Figure 6: Ratings and cumulative iconicity
+### Figure 5: Ratings and cumulative iconicity
 
@@
 plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures/fig6-panelAB_ratings.png",height=5,width=9,bg="white")
+ggsave("figures/fig5-panelAB_ratings.png",height=5,width=9,bg="white")
 
@@
 plot_grid(pA,pB,labels=c("A","B"),label_size=14,rel_widths = c(1.4,2))
-ggsave("figures/fig6-panelAB_ratings_variant1.png",height=5,width=9,bg="white")
+ggsave("figures/fig5-panelAB_ratings_variant1.png",height=5,width=9,bg="white")
 

### 2024-09-12 paper_03_exploratory_analyses.R

@@ -2,6 +2,7 @@ library(ggrepel)
 library(tidyverse)
 library(readxl)
 library(viridis)
+library(ggthemes)
 #### For some reason writing the d dataset from paper_coding_2_analysis.Rmd
 #### to utf-8 and then reading it in here as a utf-8 file didn't work :(
 
@@ -44,9 +45,10 @@ dat$category <- factor(dat$category,levels=c("Sound","Motion","Shape","Texture",
 dat |>
   ggplot(aes(x=rating_z,y=score_z,color=C_cumulative))+geom_point()+
   labs(x="rating (z)",y="guessabing accuracy (z)",colour="cumulative iconicity")+
-  facet_wrap(~category)+
-  scale_colour_viridis(option="plasma")+
-  theme_minimal()
+  scale_colour_viridis(option="plasma",begin=0.3,end=0.9) +
+  theme_tufte() +
+  facet_wrap(~category)
+ggsave("figures/fig6-z_by_domain.png",height=4,width=9,bg="white")
 
 ## Compute correlations between guesses and ratings
 

### 2024-09-12 paper_04_publication_quality_plots.R

@@ -1,0 +1,308 @@
+# publication quality figures
+
+# Packages
+list.of.packages <- c("tidyverse","readxl","extrafont",
+                      "ggthemes","ggrepel","gghalves","ggbeeswarm","viridis","cowplot")
+new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages)
+lapply(list.of.packages, require, character.only=T)
+
+font_import()
+
+
+# Data --------------------------------------------------------------------
+
+
+d <- readr::read_csv("data/ideophones_coded_guessed_rated.csv")
+
+d.ratings.full <- readr::read_csv("data/ideophones_rated.csv")
+
+
+# make a copy dp of the data in which we convert some measures to factors for easy plotting & faceting
+tofactors <- paste(c("language|category|study|group",names(d[grep('C_',names(d))])),collapse = "|")
+dp <- d
+dp[,grep(tofactors,names(dp))] <- lapply(dp[,grep(tofactors,names(dp))], as.factor)
+
+
+# Ggplot theme ------------------------------------------------------------
+
+library(extrafont)
+loadfonts(device="win")
+
+old <- theme_set(theme_tufte())
+theme_set(old)
+theme_update(text = element_text(family="Gill Sans MT"))
+
+
+# Figure 3 ----------------------------------------------------------------
+
+
+# panel A: scores by study and cumulative iconicity
+pA <- ggplot(data=dp, aes(x=study,y=score,fill=C_cumulative,colour=C_cumulative)) +
+  ylim(0,1) + theme(legend.position="none") +
+  labs(title="Guessability by study",
+       x="study",
+       y="guessability (% correct)") +
+  scale_fill_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  scale_colour_viridis(option="plasma",discrete=T,begin=0.3,end=0.9) +
+  stat_summary(fun.y=median,geom="point",size=8,shape=21,stroke=1,fill="white") +
+  geom_dotplot(stackgroups=T,dotsize=1.5,binwidth=0.01,binaxis="y",stackdir = "center") +
+  NULL
+
+pB <- ggplot(data=dp, aes(x=C_cumulative,y=score_z,colour=C_cumulative)) +
[…]

### 2024-11-17 paper_01_consistency.Rmd
de-anonymized and re-rendered to check that everything still works

@@
 title: "Coding, consistency & calibration"
-author: "[anonymized for review]"
-date: "Updated `r format(Sys.Date())`"
+author: "Mark Dingemanse, Stella Punselie, Bonnie McLean"
+date: "Rendered `r format(Sys.Date())`"
 output: github_document
@@
 list.of.packages <- c("tidyverse","readxl","writexl","irr")
-# new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
-# if(length(new.packages)) install.packages(new.packages)
+new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
+if(length(new.packages)) install.packages(new.packages)
 lapply(list.of.packages, require, character.only=T)
@@
 
-write_xlsx(d.F.inconsistent,path="data/ideophones_F_consistency_2ndround.xlsx")
+#commented out to avoid overwriting
+#write_xlsx(d.F.inconsistent,path="data/ideophones_F_consistency_2ndround.xlsx")
 

### 2024-11-17 paper_02_main_analyses.Rmd
de-anonymized and re-rendered to check that everything still works

@@
 title: "Triangulating iconicity: Coding analysis"
-author: "[anonymised for review]"
+author: "Mark Dingemanse, Stella Punselie, Bonnie McLean"
 date: "Updated `r format(Sys.Date())`"
@@
 
-Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores.
+Code notebook for a study of the relation between linguistically informed iconicity coding and experimentally collected guessability scores. Published paper is here:
+
+Punselie, S., McLean, B., & Dingemanse, M.
+(2024).
+The Anatomy of Iconicity: Cumulative Structural Analogies Underlie Objective and Subjective Measures of Iconicity.
+*Open Mind*, 8, 1191–1212.
+doi: [10.1162/opmi_a_00162](https://doi.org/10.1162/opmi_a_00162)
 
@@
 ## Data
+
 Here we load the data. For earlier processing operations, see `paper_00_data_processing.R`.
@@
 ## Figures
+
 Here we generate some descriptive stats and create the main figures for the paper. Note that Figures 1 and 2 in the paper are illustrative and their final versions are not generated in R.
@@
 ### Figure 2: Overview panel
+
 The overview of the triangulation method is created separately in CorelDraw but we use the data from the guessability studies as a canvas.
@@
 ### Coding and guessability
+
 As the visualizations already make clear there is a non-trivial positive relation between the presence of iconic mappings (in a cumulative sense) and the guessability score determined in experimental work.
@@
 
-For the fixed effects, m1 lists a 0.182 improvement in logodds for C_cumulative. According to my interpretation this means that every added iconic feature leads to a ~7% boost in proportion correct.
+For the fixed effects, m1 lists a 0.182 improvement in logodds for C_cumulative. According to my interpretation this means that every added iconic feature leads to a \~7% boost in proportion correct.
 
@@
 
-* `C_aspect`: 0.181 logodds (+7% boost)
-* `C_modality`: 0.177 logodds (+7% boost)
-* `C_length`: 0.443 logodds (+17% boost)
+-   `C_aspect`: 0.181 logodds (+7% boost)
+-   `C_modality`: 0.177 logodds (+7% boost)
+-   `C_length`: 0.443 logodds (+17% boost)
 
@@
 
-* `C_length`: `r round(m8.estimates["C_length"],2)`
-* `C_modality`: `r round(m8.estimates["C_modality"],2)`
-* `C_aspect`: `r round(m8.estimates["C_aspect"],2)`
+-   `C_length`: `r round(m8.estimates["C_length"],2)`
+-   `C_modality`: `r round(m8.estimates["C_modality"],2)`
+-   `C_aspect`: `r round(m8.estimates["C_aspect"],2)`
